### 2019-07-05 manuscript/si.Rmd

---
title: "Supporting Information"

author:

- name: \*Provisional author list\*
  affilnum: '1'

- name: Lyndon D. Estes
  affilnum: '1'
  email: lestes@clarku.edu

- name: Su Ye
  affilnum: '1'

- name: Lei Song
  affilnum: '1'

- name: Ron Eastman
  affilnum: '1'

- name: Sitian Xiong
  affilnum: '1'

- name: Tammy Woodard
  affilnum: '1'

- name: Boka Luo
  affilnum: '1'

- name: Dennis McRitchie
  affilnum: '2'

- name: Ryan Avery
  affilnum: '3'

- name: Kelly Caylor
  affilnum: '3'

- name: Stephanie Debats
  affilnum: '4'

- name: SpatialCollective
  affilnum: '5'

- name: Meridia
  affilnum: '6'

- name: Azavea
  affilnum: '7'

affiliation:

- affilnum: 1
  affil: Graduate School of Geography, Clark University, 950 Main Street, Worcester, MA 01610 USA

- affilnum: 2
  affil: Dennis and Sons, Tucson, AZ, USA

- affilnum: 3
  affil: UCSB

- affilnum: 4
  affil: Uber

- affilnum: 5
  affil: SpatialCollective

- affilnum: 5
  affil: Meridia

- affilnum: 5
  affil: Azavea

output:

  pdf_document:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    keep_tex: yes
    number_sections: yes
    template: manuscript.latex
    includes:
      in_header: header.tex

  html_document: null

  word_document: null

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: yes
spacing: doublespacing
footerdate: yes
abstract: The abstract can go either here or below
keywords: rmarkdown, reproducible science
bibliography: 
  - references.bib
  - knitcitations.bib
csl: ecology.csl

---

```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library(knitr)
library(citr)

#opts_knit$set(root.dir=normalizePath('../'))

### Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
opts_chunk$set(cache = 2, cache.path = "output/cache/")

## Plots
opts_chunk$set(fig.path = "output/figures/")

```

```{r knitcitations, echo=FALSE, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```

\singlespace

\bleft
# Methods
## Mapping platform

### Digitizing tools
To minimize the risk of topological errors, *mapper*'s polygon digitizing tools prevent drawing that results in self-intersections and overlaps between adjacent polygons. Upon submission, the PostGIS [`ST_MakeValid`](https://www.postgis.net/docs/ST_MakeValid.html) function is applied to each polygon's geometry to clean remaining topological errors upon insertion into the database.

### Accuracy assessment

### Consensus labelling
As described in the main text, the formula used for creating a consensus label is:

\begin{equation} \label{eq:main}
\mathrm{P(\theta|D)=\sum_{i=1}^{n}P(W_i|D)P(\theta|D, W_i)}
\end{equation}

Where $\theta$ represents the true cover type of a pixel (field or not field), *D* is the worker's label of that field, and *W$_i$* is an individual worker.  Looking in greater details at this equation, the first half of the righthand side of the equation, P(W$_i$|D), is the "prior" for worker *i* for the current site based on their history of scores from prior accuracy assessment assignments. The second term, P($\theta$|D, W$_i$), is the probability that the actual class of the pixel in the current assignment is the class that worker *i*'s says that it is, which is either 0 or 1. There are four possible values for this second term:

\begin{equation} \label{eq:tp}
P(\theta = field|D_i = field) = 1
\end{equation}
\begin{equation} \label{eq:fp}
P(\theta = no field|D_i = field) = 0
\end{equation}
\begin{equation} \label{eq:tn}
P(\theta = no field|D_i = no field) = 1
\end{equation}
\begin{equation} \label{eq:fn}
P(\theta = field|D_i = no field) = 0
\end{equation}

Where equations \ref{eq:tp} and \ref{eq:tn} represent true positives and negatives, respectively, and equation \ref{eq:fp} is a false positive, and equation \ref{eq:fn} is a false negative.

Coming back to the first term, the calculation of prior probability can be re-expressed as:

\begin{equation} \label{eq:prior}
\mathrm{P(W_i|D) \approx P(D|W_i)P(W_i)}
\end{equation}

Where:

\begin{equation} 
\mathrm{P(D|W_i) \propto exp\left(-\frac{1}{2}BIC_i\right)}
\end{equation}

With BIC being the Bayesian information criterion:

\begin{equation}
\mathrm{BIC = ln(n)k - 2ln(\hat{L})}
\end{equation}

In which *n* is the sample size, *k* is the number of parameters to estimate, and $\hat{L}$ is the maximum likeihood function. In this case, we are only interested in one parameter (the label that maximizes the likelihood function), thus the BIC becomes:

\begin{equation}
\mathrm{BIC \approx -2ln(\hat{L}) = 2ln(p(D|\hat{\theta}, W))}
\end{equation}

After rearranging, we have:

\begin{equation}
\mathrm{P(D|W_i) \propto p(D|\hat{\theta}, W_i))}
\end{equation}

Which is the worker maximum likelihood, which can be computed as:

\begin{equation} \label{eq:pacc1}
\mathrm{P(\theta = field|\hat{\theta}, M_I) = P(D = field|\theta = field, M_I) = \frac{1}{m}\left(\sum_{j}^{m}\frac{tp_j}{tp_j + fn_j} \right)}
\end{equation}

\begin{equation} \label{eq:pacc2}
\mathrm{P(\theta = no field|\hat{\theta}, M_I) = P(D = no field|\theta = no field, M_I) = \frac{1}{m}\left(\sum_{j}^{m}\frac{tn_j}{tn_j + fp_j} \right)}
\end{equation}

Equations \ref{eq:pacc1} and \ref{eq:pacc2} are producer's accuracies, thus the maximum worker likelihood is equivalent to the worker's average producer's accuracy.

<!-- \ref{eq:prior} represents "mapper likelihood" for the current map: -->

The other component of equation \ref{eq:prior}, P(W$_i$), is the worker's average score over *m* accuracy assessment assignment:  
\begin{equation}
\mathrm{P(W_i) \propto \frac{1}{m}\sum_{j=1}^{m}score_j}
\end{equation}

Thus equation \ref{eq:prior} uses two measures of worker accuracy, 1) their overall average accuracy score multipled by 2) their average producer's accuracy to create a *weight* for their individual maps for the given site. Equation \ref{eq:main} becomes:

\begin{equation} 
\mathrm{P(\theta|D)=\frac{\sum_{i=1}^{n}weight_iP(\theta|D, W_i)}{\sum_{i=1}^{n}weight_i}}
\end{equation}

With $\mathrm{P(\theta|D, W_i)}$ being either 0 or 1. In labelling, if the consensus result for a pixel is: $\mathrm{P(\theta = field|D)}$ > 0.5, then we assign that pixel to the field category, otherwise to the no field category.

#### Example

To provide an example of this approach in practice, we'll imagine two workers A and B, each with the following histories:

```{r, echo = FALSE}
library(magrittr)
sc <- tibble::tibble(Worker = LETTERS[1:2], `Prod. Acc. (Field)` = c(0.8, 0.62),
                     `Prod. Acc (no field)` = c(0.81, 0.61), 
                     Score = c(0.75, 0.60))
wtA <- sc[1, ]$Score * sc[1, ]$`Prod. Acc. (Field)`
wtB <- sc[2, ]$Score * sc[2, ]$`Prod. Acc (no field)`

# sc
knitr::kable(sc, align = c("r", "c", "c"), format = "pandoc")
```

In this scenario, worker A thinks that the given pixel falls within a field, and worker B thinks it is not a field.  First, we calculate the weights for each worker:

$$
Weight_A = score_A * PA_A(field) = P(W_A)P(D=field|W_A) = `r sc[1, 2]` * `r sc[1, 4]` = `r wtA`
$$

$$
Weight_B = score_B * PA_B(field) = P(W_B)P(D=no field|W_B) = `r sc[2, 3]` * `r sc[2, 4]` = `r wtB`
$$

And then we plug these weights into the full equation:

$$
\mathrm{P(\theta|D)=\frac{\sum_{i=1}^{n}weight_iP(L = field|D, W_i)}{\sum_{i=1}^{n}weight_i}} = \frac{`r wtA` * 1 + `r wtB` * 0}{`r wtA` + `r wtB`} = `r round((wtA * 1 + wtB * 0) / (wtA + wtB), 3)`
$$

Since `r round((wtA * 1 + wtB * 0) / (wtA + wtB), 3)` > 0.5, we label the particular pixel a field.

# References
\singlespace

<div id = "refs"></div>

\eleft

\clearpage

```{r sessioninfo, echo = FALSE, eval = FALSE}
# set eval = FALSE if you don't want this info (useful for reproducibility) to appear 
sessionInfo()
```