### 2021-03-02 inst/manuscript/si.Rmd
Edited Results and SI

---
title: "Supporting Information: High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales"

output:
  pdf_document:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    keep_tex: yes
    number_sections: yes
    template: manuscript.latex
    includes:
      in_header: header.tex

  html_document: null

  word_document: null

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: false
spacing: doublespacing
# abstract: The abstract can go either here or below
keywords: rmarkdown, reproducible science
bibliography: 
  - references.bib
  - knitcitations.bib
csl: ecology.csl

---

<!-- Provides custom captions for figures in supplement -->
\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

```{r setup, include=FALSE, cache=FALSE, message=FALSE}
library(knitr)
library(citr)
library(tidyverse)
library(sf)
library(activemapper)

#opts_knit$set(root.dir=normalizePath('../'))

### Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
opts_chunk$set(cache = 2, cache.path = "output/cache/")

## Plots
opts_chunk$set(fig.path = "output/figures/")

knitr::opts_chunk$set(fig.pos = "!ht", out.extra = "")

```

```{r knitcitations, echo=FALSE, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```

\beginsupplement
\singlespace

\bleft

# Methods
## Assessing quality of composites
The rubric presented in Table \ref{tab:imqcriteria} was used to assess the quality of the seasonal image composites. The imagery was evaluated by examining their Raster Foundry overlays within a `labeller` instance set up for the purpose.

```{r imqcriteria, echo=FALSE, message=FALSE}
library(kableExtra)
image_qual_criteria <- tibble(
  `Quality dimension` = c(
    "Percent affected by residual cloud", "Percent affected by cloud shadow", 
    "Number of visible scene boundaries", "Percent blurred"
  ),
  `3 pts` = c("<1%", "<1%", "None", "None"), 
  `2 pts` = c("1-5%", "1-5%", "1", "<20%"), 
  `1 pts` = c("5-10%", "5-10%", "2-3", "20-50%"),
  `0 pts` = c(">10%", ">10%", ">3", ">50%")
) 
cap <- glue::glue("Four dimensions used to assess the assess the quality
                  of the temporally composited image tiles, including the
                  criteria used to award points for scoring each dimension.")
kable(image_qual_criteria, format = "latex", booktabs = "T", caption = cap) %>% 
  kable_styling(full_width = TRUE, latex_options = c("hold_position")) %>% 
  column_spec(1, width = "8cm")
```

## Mapping platform
### Digitizing tools
To minimize the risk of topological errors, *mapper*'s polygon digitizing tools prevent drawing that results in self-intersections and overlaps between adjacent polygons. Upon submission, the PostGIS [`ST_MakeValid`](https://www.postgis.net/docs/ST_MakeValid.html) function is applied to each polygon's geometry to clean remaining topological errors upon insertion into the database.

### Mapping rules
Labellers were asked to follow a particular set of rules when digitizing field boundaries. The following rules are reproduced (with some modification to wording) from a [website](https://github.com/agroimpacts/mappingrules) that we developed to host the rules.

#### Rules for interpreting and mapping imagery

- **Rule 1**: Map the fields that are visible in the Planet imagery, not the fields that you see in the basemap.

- **Rule 2**: Please map the fields as they appear in the off-season Planet images. If it is too hard to make these out clearly, the next choice is to map them in growing season imagery.

- **Rule 3**: Use all available images (Planet scenes and base map) to help you interpret what is and is not a field. Toggling back and forth between the different image sources can help you identify and classify the field boundaries.

- **Rule 4**: Map fields that intersect the mapping target, making sure to finish the entire boundary.

#### Rules for classifying different types of agriculture

- **Rule 5**: Choose the annual croplands category if the fields looks like they contain crops that are planted and harvested during a single year (or occasionally after a few seasons).

- **Rule 6**: Fields that look like they are under tree crops, such as orchards, woodlots, or other planted forest types are agriculture, but we will not map these in the current campaign. Therefore, do not draw polygons around these types of fields.

- **Rule 7**: Choose the fallow category if the fields looks like it was an annual cropland that is somewhat overgrown, and hasn't been planted in a few seasons. [Note: this category was eventually not used, as such fields were not mapped]

- **Rule 8**: Choose the field, but unsure of category if you can't tell whether the field is annual cropland, a tree crop, or fallow. Try hard to avoid this category. [Note: this category was also not used]

- **Rule 9**: Choose the unsure if agricultural field or not category if you think the field might be agriculture (whether annual cropland, tree crops, or fallow), but it could also be non-agricultural land. Try hard to avoid this category also. [Note: this category was also not used]

- **Rule 10**: Don't map any polygons if there is no agricultural to map, or if the fields look like they are tree crops (see updated Rule 6).

### Assessing label accuracy
For each accuracy assessment assignment, the labeller's maps are scored against  training reference polygons digitized by one of the map supervisors (e.g. Estes or Ye). A proportion of training reference sites were of the non-cropland class and thus had no corresponding polygons.

As described in the main text, the score for a particular assignment *i* is calculated as the weighted sum of five accuracy metrics:

$$ score_i=\beta_0\mathrm{I}+\beta_1\mathrm{O}+\beta_2\mathrm{F}+\beta_3\mathrm{E}+\beta_4\mathrm{C}$$

Where $\beta_{0-5}$ are the weights assigned to each accuracy metric that sum to 1. For the current production version, $\beta_{0-5}$ were assigned as 0.4, 0.2, 0.2, 0.1, and 0.1.

*I* is "inside the box" accuracy which is defined as a proportion of the area correctly mapped within a 0.005 degree resolution grid (*I$_c$*) over total "inside the box" area for this grid (*I$_t$*).

$$ I=\frac{I_c}{I_t}$$

*O* is "outside the box" accuracy which refers to a proportion of the field area correctly mapped outside the grid over total "outside the box" region (*O$_t$*, the region within the bounding box of the workers' polygons but not within 0.005 degree resolution grid).

$$ O=\frac{O_c}{O_t}$$

*F* is the fragmentation accuracy which is defined as a proportion of matched polygon number (*N$_m$*, the number of the workers' polygons that has at least 50% of its region overlapped with a reference polygon) over total workers' polygon number (*N$_t$*).

$$ F=\frac{N_m}{N_t}$$

*E* is the average edge accuracy for all pairs of matched workers and reference polygons; the edge accuracy for a single pair is defined as the length of 'correctly mapped edges' (*L$_c$*, the partial boundary of a workers' polygon that are within a three-pixel buffer region of the matched reference polygon boundary) over the total boundary length of its matched reference polygon (*L$_t$*).

$$ E=\frac{L_c}{L_t}$$

*C* is the categorical accuracy, i.e., a proportion of the area that has been correctly labeled with field category within intersected regions between worker's and reference polygons (*T$_c$*) over the total intersecting area (*T$_t$*).

$$ C=\frac{T_c}{T_t}$$

### Consensus labelling
As described in the main text, the formula used for creating a consensus label is:

\begin{equation} \label{eq:main}
\mathrm{P(\theta|D)=\sum_{i=1}^{n}P(W_i|D)P(\theta|D, W_i)}
\end{equation}

Where $\theta$ represents the true cover type of a pixel (field or not field), *D* is the worker's label of that field, and *W$_i$* is an individual worker.  Looking in greater details at this equation, the first half of the righthand side of the equation, P(W$_i$|D), is the "prior" for worker *i* for the current site based on their history of scores from accuracy assessment assignments. The second term, P($\theta$|D, W$_i$), is the probability that the actual class of the pixel in the current assignment is the class that worker *i*'s says that it is, which is either 0 or 1. There are four possible values for this second term:

\begin{equation} \label{eq:tp}
P(\theta = field|D_i = field) = 1
\end{equation}
\begin{equation} \label{eq:fp}
P(\theta = no field|D_i = field) = 0
\end{equation}
\begin{equation} \label{eq:tn}
P(\theta = no field|D_i = no field) = 1
\end{equation}
\begin{equation} \label{eq:fn}
P(\theta = field|D_i = no field) = 0
\end{equation}

Where equations \ref{eq:tp} and \ref{eq:tn} represent true positives and negatives, respectively, and equation \ref{eq:fp} is a false positive, and equation \ref{eq:fn} is a false negative.

Coming back to the first term, the calculation of prior probability can be re-expressed as:

\begin{equation} \label{eq:prior}
\mathrm{P(W_i|D) \approx P(D|W_i)P(W_i)}
\end{equation}

Where:

\begin{equation} 
\mathrm{P(D|W_i) \propto exp\left(-\frac{1}{2}BIC_i\right)}
\end{equation}

With BIC being the Bayesian information criterion:

\begin{equation}
\mathrm{BIC = ln(n)k - 2ln(\hat{L})}
\end{equation}

In which *n* is the sample size, *k* is the number of parameters to estimate, and $\hat{L}$ is the maximum likeihood function. In this case, we are only interested in one parameter (the label that maximizes the likelihood function), thus the BIC becomes:

\begin{equation}
\mathrm{BIC \approx -2ln(\hat{L}) = 2ln(p(D|\hat{\theta}, W))}
\end{equation}

After rearranging, we have:

\begin{equation}
\mathrm{P(D|W_i) \propto p(D|\hat{\theta}, W_i))}
\end{equation}

Which is the worker maximum likelihood, which can be computed as:

\begin{equation} \label{eq:pacc1}
\mathrm{P(\theta = field|\hat{\theta}, M_I) = P(D = field|\theta = field, M_I) = \frac{1}{m}\left(\sum_{j}^{m}\frac{tp_j}{tp_j + fn_j} \right)}
\end{equation}

\begin{equation} \label{eq:pacc2}
\mathrm{P(\theta = no field|\hat{\theta}, M_I) = P(D = no field|\theta = no field, M_I) = \frac{1}{m}\left(\sum_{j}^{m}\frac{tn_j}{tn_j + fp_j} \right)}
\end{equation}

Equations \ref{eq:pacc1} and \ref{eq:pacc2} are producer's accuracies, thus the maximum worker likelihood is equivalent to the worker's average producer's accuracy.

<!-- \ref{eq:prior} represents "mapper likelihood" for the current map: -->

The other component of equation \ref{eq:prior}, P(W$_i$), is the worker's average score over *m* accuracy assessment assignment:  
\begin{equation} 
\mathrm{P(W_i) \propto \frac{1}{m}\sum_{j=1}^{m}score_j}
\end{equation}

Thus equation \ref{eq:prior} uses two measures of worker accuracy, 1) their overall average accuracy score multipled by 2) their average producer's accuracy to create a *weight* for their individual maps for the given site. Equation \ref{eq:main} becomes:

\begin{equation} 
\mathrm{P(\theta|D)=\frac{\sum_{i=1}^{n}weight_iP(\theta|D, W_i)}{\sum_{i=1}^{n}weight_i}}
\end{equation}

With $\mathrm{P(\theta|D, W_i)}$ being either 0 or 1. In labelling, if the consensus result for a pixel is: $\mathrm{P(\theta = field|D)}$ > 0.5, then we assign that pixel to the field category, otherwise to the no field category.

After creating the consensus label, the degree of confidence in the resulting label value is measured by Bayesian Risk:

\begin{equation} \label{eq:pixelrisk}
\mathrm{r=C(1 - L) + (1 - C)L}
\end{equation}

Where *C* is the consensus probability that a given pixel is a field ($\mathrm{P(\theta = field|D)}$), and *L* is the consensus label (i.e. non-field if *C* < 0.5, field if *C* > 0.5) for that pixel. The risk values across the entire sample site can be processed in two ways to provide useful information about the confidence in the consensus label for that site. The first is a simple average of all risk values in the site, where the slope of the risk varies depending on whether *L* is a field or not a field (Figure \ref{fig:riskcurve}). The closer to 0 the lower the risk that the *L* is mislabelled, while values approaching 1 indicate increasing risk of mislabelling. A second approach is to calculate the proportion of pixels having risk values that exceed a certain threshold.

```{r riskcurve, echo = FALSE, fig.align='center', out.width="70%", message=FALSE, warning=FALSE, fig.cap="Bayesian risk values (Y-axis) for consensus values (X axis) ranging from 0 to 1 (0 indicates no consensus that a pixel falls into the field class, 1 means complete consensus) for field and non-field consensus labels."}
knitr::include_graphics('figures/si_label_risk.png')
```

Accuracy assessment and consensus label generation is performed using an `R` package `rmapaccuracy` that is internal to the labelling platform. The package relies on spatial data classes and operations provided by the `sf` [@pebesmaSimpleFeaturesStandardized2018] and `raster` [@hijmansRasterGeographicData2020] packages.

#### Example
To provide an example of this approach in practice, we'll imagine two workers A and B, each with the following histories:

```{r, echo = FALSE}
library(magrittr)
sc <- tibble::tibble(Worker = LETTERS[1:2], `Prod. Acc. (Field)` = c(0.8, 0.62),
                     `Prod. Acc (no field)` = c(0.81, 0.61), 
                     Score = c(0.75, 0.60))
wtA <- sc[1, ]$Score * sc[1, ]$`Prod. Acc. (Field)`
wtB <- sc[2, ]$Score * sc[2, ]$`Prod. Acc (no field)`

# sc
knitr::kable(sc, align = c("r", "c", "c"), format = "pandoc")
```

In this scenario, worker A thinks that the given pixel falls within a field, and worker B thinks it is not a field.  First, we calculate the weights for each worker:

\clearpage
$$
Weight_A = score_A * PA_A(field) = P(W_A)P(D=field|W_A) = `r sc[1, 2]` * `r sc[1, 4]` = `r wtA`
$$

$$
Weight_B = score_B * PA_B(field) = P(W_B)P(D=no field|W_B) = `r sc[2, 3]` * `r sc[2, 4]` = `r wtB`
$$

And then we plug these weights into the full equation:

$$
\mathrm{P(\theta|D)=\frac{\sum_{i=1}^{n}weight_iP(L = field|D, W_i)}{\sum_{i=1}^{n}weight_i}} = \frac{`r wtA` * 1 + `r wtB` * 0}{`r wtA` + `r wtB`} = `r round((wtA * 1 + wtB * 0) / (wtA + wtB), 3)`
$$

Since `r rj <- round((wtA * 1 + wtB * 0) / (wtA + wtB), 3); rj` > 0.5, we label the particular pixel a field.

Using equation \ref{eq:pixelrisk}, the corresponding risk associated with this particular pixel's label is thus `r rj * (1 - 1) + (1 - rj) * 1` (i.e., 1 - `r rj <- round((wtA * 1 + wtB * 0) / (wtA + wtB), 3); rj`).

### Accuracy assessment
We designed and implemented a map accuracy assessment protocol following procedures summarized by @StehmanKeyissuesrigorous2019. This entailed the creation of a map reference sample, which first entailed designing a sample, and then designing how the sample response would be collected.

#### Map reference sample design
We employed a stratified design for collecting the map reference sample, using the segmented field boundaries to define the strata for cropland/non-cropland. To create the sample, we first extracted the centroids for each of the sample grid cells in Ghana. We then intersected the centroid points with the field segments, assigned a class of 1 (cropland) where points intersected a field, and 0 where they didn't (non-cropland). We removed from this set of points all those that corresponded to model training, validation, or training reference sites, and extracted a random sample from both the cropland and non-cropland points. To determine the sample size of each, we specified a desired confidence interval using the following formula [@StehmanKeyissuesrigorous2019]:

$$n = \frac{z^2p(1-p)}{d^2}$$

Where p is the estimated probability (or mapped class accuracy), and d is the size of the margin of error (1/2 the confidence interval). We selected a d value of 0.03 and assumed that the user's accuracy of the field class would be 0.75 and that of the non-cropland class would be 0.8, returning sample sizes of 800 and 683, respectively. The distribution of the resulting map reference sample is shown in Figure \ref{fig:refsample}.

```{r refsample, echo = FALSE, fig.align='center', out.width="80%", message=FALSE, warning=FALSE, fig.cap="Distribution of the selected map reference sample for the cropland and non-cropland class. "}
knitr::include_graphics('figures/si_map_reference_sample.png')
```

#### Response design
We collected the map reference sample on a separate instance of the labelling platform set up for the purpose. For the sampling unit, we selected a rectangular polygon of ~0.1 ha (0.0002866$^\circ$ resolution). This polygon was centered on the centroid of each grid cell selected for map reference sample, and was presented as the target grid in the labelling platform. Four classes were established for the validation: **cropland**, **non-cropland**, **uncertain but likely cropland**, and **uncertain but likely non-cropland**. The latter two classes were designed to capture information related to swidden dynamics, following the rationale that uncertainty and time since last cropping are likely to be positively correlated. This uncertainty also captures information about the inherent difficulty of the mapping task. Samples were collected by visually interpreting the overlays of PlanetScope composites, following the same interpretation protocols used by the labelling team, with the exception that the polygons placed were square and of 0.0002866$^\circ$ resolution.

Two supervisors (Ye and Estes) who were not involved in labelling developed the map reference sample evaluated the PlanetScope composites at each site to determine its class membership. The following set of rules were followed in collecting the sample:

1. When determining the class corresponding with the initial location of the target grid, if:
  - More than half of the target falls within what appears to be a clear arable crop field, then classify it as **cropland**; 
  - More than half falls in what is clearly not a field, then classify it as **non-cropland**; 
  - More than half falls in a location where it is harder to tell whether it is cropland or non-cropland, determine whether it is more likely a crop field or not a crop field, and then assign either **uncertain but likely cropland**, or **uncertain but likely non-cropland**.

2. After determining the class, if
  - The target polygon is contained entirely within a single clear class, then simply digitize a point within the center of the target box, assign the appropriate class label, and complete the assignment;
  - Digitize a square polygon exactly aligned with the initial target, choose the correct class label, and then move the new polygon to the nearest location where it can be contained entirely by the assigned class.

After collecting the sample, the geometries were further refined by converting the points (sites were the target didn't have to be shifted) to polygons with the same  0.0002866$^\circ$ resolution, and then the full set of map reference polygons was used to extract the classified values from both the cropland probability and vectorized field boundary maps.

\clearpage
# Results
## Image quality
```{r, echo=FALSE, message=FALSE, warning=FALSE}
data("image_quality")
wqual_diff <- image_quality$img %>% filter(kml_type == "F") %>%
  group_by(name, season, worker_id) %>%
  summarize(tile = min(tile), score = sum(score) / 12) %>% 
  ungroup() %>% 
  pivot_wider(names_from = worker_id, values_from = score,
              names_prefix = "w") %>%
  select(tile, season, w92, w99) %>% 
  mutate(d = w92 - w99, 
         absd = abs(d))
```

The image quality assessment was conducted by two separate observers (L. Song and Q. Zhang). Each observer assessed the composites for each season at all 50 of the randomly selected tiles (see main text) following the defined criteria (Table \ref{tab:imqcriteria}). To calculate the score for a given tile for each season, we summed the ranking across the four categories for each observer, rescaled the values, and then calculated the mean tile score across both observers by season. The mean difference between the two observers was `r d <- round(mean(wqual_diff$d), 3); d` (sd = `r absd <- round(sd(wqual_diff$d), 3); absd`) and the mean absolute difference was `r round(mean(wqual_diff$absd), 3)` (sd = `r round(sd(wqual_diff$absd), 3)`), thus one observer scored tiles about `r abs(round(d * 100))` percent lower than the other observer, on average.

## Active learning process
### Labelling
The distributions of the initial 500 sites used to train the RandomForest model for each AOI are shown in Figure \ref{fig:trainval}. In AOI 3, the initial active learning cycle resulted in low accuracy because the northern part of the AOI shows low contrast between fields and the surrounding vegetation in the dry season. Training the model with the initial 500 samples resulted in large commission errors in this part of the AOI, thus we ran a second active learning cycle that began with an inital random draw of 300 training sites confined to this AOI (blue points in Figure \ref{fig:trainval}A)
```{r trainval, echo = FALSE, out.width="80%", fig.cap="The distribution of A) initial randomly selected training sites, including 300 points selected within AOI 3 to retrain a second active learning run for this region, as well as the locations of training reference sites, and B) validation points and training sites selected during each active learning iteration.", fig.align='center', message=FALSE, results='hold'}
knitr::include_graphics('figures/si_training_validation_pts.png')
```

The distribution of data collection effort and scores from training reference sites by labelling team are shown in Figure \ref{fig:trainqual}. These results include those from re-labelling the initial training in Cluster 2 (AOIs 7-9, 12, 15). This was done because we discovered a small spatial offset in the original image composites, which we corrected by reprocessing the images. We replaced the image overlays in the instance for central-southeast \ref{fig:trainval}A), and the labelling team mapped these sites a second time during the production run in late 2019. The reprocessed labels were used to initially train the model for AOIs 9, 12, and 15.

```{r trainqual, echo = FALSE, out.width="60%", fig.cap="The A) distributions of scores at training reference sites for each labeller (means indicated by X in boxplots), and B) the number of training and training reference sites completed by each labeller. Labellers' identities are anonymized.", fig.align='center', message=FALSE, results='hold'}
knitr::include_graphics('figures/si_training_qual_n.png')
```

The average Bayesian Risk of each training and validation site is shown in Figure \ref{fig:labelrisk}, as well as the distribution of risk values per AOI and the initial training clusters (Figure \ref{fig:labelriskhist}). The three initial clusters include the second mapping of Cluster 2.

```{r labelrisk, echo = FALSE, out.width="80%", fig.cap="The average Bayes Risk of each training and validation site in Ghana.", fig.align='center', message=FALSE, results='hold'}
knitr::include_graphics('figures/si_label_risk_map.png')
```

```{r labelriskhist, echo = FALSE, out.width="45%", fig.cap="The distribution of Bayesian Risk values for training (left column) and validation (right column) sites in each AOI, with the average value indicated by vertical lines. The first three rows (panel titles beginning with 'Cl') indicate distributions of Bayes Risk in the initial training clusters, including the first ('Cl2a') and second ('Cl2b') mappings of Cluster 2.", fig.align='center', message=FALSE, results='hold'}
knitr::include_graphics('figures/si_label_risk_hists.png')
```

### Validation
The differences in accuracy, AUC, and F1 between the active learning process and the random retraining at each iteration for AOIs 1, 8, and 15 (Figure \ref{fig:randomvactive}). Small differences due to random variations in the RandomForest models are the reason for the non-zero differences at iteration 0, when both models were trained with the same level set.   
```{r randomvactive, echo = FALSE, out.width="90%", fig.cap="The percent difference in performance metrics per iteration for AOIs 1, 8, and 15 (grey lines with numbers indicating AOI; the black line indicates average difference across the three AOIs) when comparing models trained using active learning versus those trained using randomly selected sites. Postive percentages indicate superior performance by active learning, negative percentages the inverse.", fig.align='center', message=FALSE, results='hold'}
knitr::include_graphics('figures/si_random_vs_active.png')
```

### The impact of training data error
Probability images resulting from Random Forest models trained with labels generated by 1) consensus and the 2) most and 3) least accurate worker to digitize each training site are illustrated in Figure \ref{fig:labelstrategy}. These images were created for a single tile in AOI 1.

```{r labelstrategy, echo = FALSE, out.width="100%", fig.cap="Cropland probability images produced by RandomForest models trained with A) consensus labels, B) the most accurate worker's labels, and C) the least accurate worker's labels.", fig.align='center', message=FALSE}
knitr::include_graphics('figures/si_label_strategy_probs.png')
```

The association between the average label risk per AOI and several model performance metrics, assessed using Spearman Rank Correlation, is shown in Table \ref{tab:labelrisk_v_metrics}.

```{r labelrisk_v_metrics, echo=FALSE}
data("label_risk_metrics")

cor_fun <- function(df) {
  cor.test(df$Risk, df$Score, method = "spearman") %>% broom::tidy()
}

lr_mets <- unique(label_risk_metrics$Metric)
risk_corr <- lapply(lr_mets[lr_mets != "TSS"], function(x) {
  dat <- label_risk_metrics %>% filter(Metric == x)
  cor_fun(dat) %>% mutate(Metric = x)
}) %>% do.call(rbind, .) %>% 
  select(estimate, Metric) %>% 
  rename(r = estimate)
nms <- risk_corr$Metric
risk_corr <- round(t(risk_corr[, 1]), 3)
colnames(risk_corr) <- nms

cap <- glue::glue("Spearman's Rank correlation between the average label risk per AOI and a variety of model performance metrics.")
kable(risk_corr, format = "latex", booktabs = "T", caption = cap) %>% 
  kable_styling(full_width = TRUE, latex_options = c("hold_position")) #%>% 
  # column_spec(1, width = "8cm")

# tp <- 10
# fp <- 2
# tn <- 20
# fn <- 3
# tss <- (tp * tn - fp * fn) / ((tp + fn) * (fp + tn))
# mcc <- (tp * tn - fp * fn) / sqrt((tp + fp) * (tp + fn) * (tn + fp) * (tn + fn))

```

## Map accuracy

### Reference label uncertainty

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# reference labels, for total count and uncertainty of class categories
ref_labels <- st_read(
  system.file("extdata/map_reference_labels.geojson", 
              package = "activemapper"), quiet = TRUE
)
class_ct <- as_tibble(ref_labels) %>% group_by(category) %>% count()
ref_ct <- as_tibble(ref_labels) %>% group_by(worker_id) %>% count() %>% pull(n)

crp_ct <- class_ct %>% filter(category %in% c("annualcropland", "unsure1"))
noncrp_ct <- class_ct %>% filter(category %in% c("noncropland", "unsure2"))

# observer agreement/correlation
ref_labels_joint <- st_read(
  system.file("extdata/map_reference_labels_joint.geojson", 
              package = "activemapper"), quiet = TRUE
)
joint_classes <- ref_labels_joint %>% as_tibble() %>% 
  select(name, category, worker_id) %>% 
  arrange(name) %>% 
  mutate(class = case_when(
    category == "annualcropland" | category == "unsure1" ~ 1, 
    category == "noncropland" | category == "unsure2" ~ 0 
  ))
class_mat <- joint_classes %>% dplyr::select(name, class, worker_id) %>% 
  tidyr::pivot_wider(names_from = worker_id, values_from = class) %>% 
  dplyr::select(-name) %>% as.matrix()

usertab <- table("User1" = class_mat[, 1], "User2" = class_mat[, 2])

agreement <- round(sum(diag(usertab)) / sum(usertab) * 100, 1)
observer_r <- cor.test(class_mat[, 1], class_mat[, 2], method = "spearman")

```

The size of the collected validation sample was `r sum(ref_ct)`, with `r ref_ct[2]` samples collected and interpreted by one observer (Su Ye) and `r ref_ct[1]` collected by a second observer (Estes). To evaluate the uncertainty inherent in defining the map reference labels, the pair mapped `r sum(usertab)` common sites, showing an overall level of agreement of `r agreement`%, and a Spearman Rank Correlation of `r round(unname(observer_r$estimate), 2)`.

Although this overlap between observers was limited, the map reference classification scheme provided two additional measures of uncertainty, which were classes defined as "unsure but most likely a field" or "unsure but most likely not a field", which the reference labeller would choose when they could not with high confidence state whether the site was either cropland or not cropland. Of the `r sum(crp_ct$n)` map reference samples that were identified as cropland, `r round(crp_ct$n[2] / sum(crp_ct$n) * 100, 1)`% fell into the lower confidence category, as did `r round(noncrp_ct$n[2] / sum(noncrp_ct$n) * 100, 1)`% of the `r sum(noncrp_ct$n)` non-cropland samples. Across both classes, the reference labellers had lower confidence in `r round(sum(crp_ct$n[2], noncrp_ct$n[2]) / sum(ref_ct) * 100, 1)` of sites labelled.

### Categorical accuracy
In addition to the map accuracies and area estimates calculated per AOI zone (reported in main text; see Figure \ref{fig:aoizones}A), the accuracies were also assessed within several different groupings of agroecozones (Figures \ref{fig:aoizones}B and Table \ref{tab:aezaccuracy}).

```{r aoizones, echo = FALSE, out.width="100%", fig.cap="Four zones used to combine the A) AOIs and B) agroecozones (B), in order to assess sub-national mapping accuracy. In B, the ID number of Ghana's individual agroecozones is shown in grey: 1 = Coastal savanna; 2 = Wet evergreen; 3 = Moist evergreen; 4 = Deciduous forest; 5 = Transitional zone; 6 = Guinea savanna; 7 = Sudan savanna; 8 (not shown) = Volta Lake.", fig.align='center', message=FALSE, results='hold'}
knitr::include_graphics('figures/si_aoi_zones.png')
```

\begin{table}
\caption{Map accuracies and adjusted area estimates for the ~3 m pixel-wise classifications (based on RandomForest predictions). Results are provided for four different groupings of Ghana's 8 agroecozones zones (Zone 1 = Coastal savanna; Zone 2 = Wet evergreen, Moist evergreen, and Deciduous forest; Zone 3 = Transitional zone; Zone 4 = Guinea savanna and Sudan savanna) plus the entire country. The error matrix (with reference values in columns) provides the areal percentage for each cell, and the producer's (P), user's (U), and overall (O) map accuracies and their margins of error (in parenthesis) are provided, as well as the sample-adjusted area estimates (in km$^{2}$) and margins of error.}
\includegraphics[width = 18cm]{figures/si_aez_accuracies.png}
\label{tab:aezaccuracy}
\end{table}

### Field area and number
To mean, median, and distributions of the average area of segmented field boundaries over the 100 validation sites in each AOI are compared to the areas of the polygons digitized by the most accurate labeller over the same sites in Figure \ref{fig:areavalidation}. The same statistics for average number of segments versus average number of labelled polygons across validation sites in each AOI are shown in Figure \ref{fig:numbervalidation}.      
```{r areavalidation, echo = FALSE, out.width="100%", fig.cap="The distributions of the average areas (in hectares) of segmented field boundaries (shown in blue) at the 100 validation sites per AOI, compared with the average areas of field boundaries digitized (shown in red) by the most accurate worker to label each site. Vertical lines indicate the mean and median of each distribution.", fig.align='center', message=FALSE, results='hold'}
knitr::include_graphics('figures/si_validation_stats_fha2.png')
```

```{r numbervalidation, echo = FALSE, out.width="90%", fig.cap="The distributions of the average number of segmented field boundaries (shown in blue) at the 100 validation sites per AOI, compared with the average number of digitized polygons (shown in red) by the most accurate worker to label each site. The mean and median of each distribution is shown. Vertical lines indicate the mean and median of each distribution.", fig.align='center', message=FALSE, results='hold'}
knitr::include_graphics('figures/si_validation_stats_fn2.png')
```

## Cropland characteristics
### Field area and number

To examine average field sizes, and the total number, the mean segment size per 0.05 degree tiles was calculated and mapped, as well as the total number of fields per tile (Figure \ref{fig:fieldsizes}).

```{r fieldsizes, echo = FALSE, out.width="100%", fig.cap="The A) average sizes of fields and B) total number of fields in each 0.05 degree tile, as calculated from the vectorized field boundaries.", fig.align='center', message=FALSE, results='hold'}
knitr::include_graphics('figures/si_fsize_n_05.png')
```

\clearpage
# References
\singlespace

<div id = "refs"></div>

\eleft

\clearpage

```{r sessioninfo, echo = FALSE, eval = FALSE}
# set eval = FALSE if you don't want this info (useful for reproducibility) to appear 
sessionInfo()
```