### 2021-03-12 inst/manuscript/manuscript.tex
Updating raw scripts and package vignette. Added tex files

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Original default rstudio/pandoc latex file
%% upated by @jhollist 09/15/2014
%% inspired by @cboetting https://github.com/cboettig/template and
%% @rmflight blog posts:
%% http://rmflight.github.io/posts/2014/07/analyses_as_packages.html
%% http://rmflight.github.io/posts/2014/07/vignetteAnalysis.html).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{float}  % added 9/3/2021 LDE
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Changes borrowed from @cboettig, added by @jhollist
% A modified page layout
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
\usepackage{lineno} % add
  \linenumbers % turns line numbering on
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%Packages and layout changes by @jhollist 09/15/2014
\usepackage{ragged2e}
\usepackage[font=normalsize]{caption}
  \usepackage[doublespacing]{setspace}
\usepackage{parskip}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
  \rfoot{\today}
\lfoot{\thepage}
%%Changed default abstract width and added lines
\renewenvironment{abstract}{
  \hfill\begin{minipage}{1\textwidth}
  \rule{\textwidth}{1pt}\vspace{5pt}
  \normalsize
  \begin{justify}
  \bfseries\abstractname\vspace{5pt}
  \end{justify}}
  {\par\noindent\rule{\textwidth}{1pt}\end{minipage}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Added this on 28/2/2021 to deal with CSLReferences issue:
%% >> LaTeX Error: Environment CSLReferences undefined.
%% Occurs after pandoc upgraded, per https://github.com/mpark/wg21/issues/54

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{High resolution, annual maps of the characteristics of
smallholder-dominated croplands at national scales}
\author{
Lyndon D. Estes
Su Ye
Lei Song
Boka Luo
J. Ronald Eastman
Zhenhua Meng
Qi Zhang
Dennis McRitchie
Stephanie R. Debats
Justus Muhando
Angeline H. Amukoa
Brian W. Kaloo
Jackson Makuru
Ben K. Mbatia
Isaac M. Muasa
Julius Mucha
Adelide M. Mugami
Judith M. Mugami
Francis W. Muinde
Fredrick M. Mwawaza
Jeff Ochieng
Charles J. Oduol
Purent Oduor
Thuo Wanjiku
Joseph G. Wanyoike
Ryan B. Avery
Kelly K. Caylor
}
\date{}
% Allowing for landscape pages
\usepackage{lscape}
\usepackage{graphicx}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

% Left justification of the text: see https://www.sharelatex.com/learn/Text_alignment
% \usepackage[document]{ragged2e} % already in the latex template
\newcommand{\bleft}{\begin{flushleft}}
\newcommand{\eleft}{\end{flushleft}}
\usepackage{flafter}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

%%Fix tightlist error: https://stackoverflow.com/questions/40438037/tightlist-error-using-pandoc-with-markdown
%%Added 2018-03-26
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
%%%

\begin{document}
%%Edited by @jhollist 09/15/2014
%%Adds title from YAML
\begin{singlespace}
\begin{center}
\huge High resolution, annual maps of the characteristics of
smallholder-dominated croplands at national scales
\end{center}
%%Adds subtitle from YAML
%%Adds Author, correspond email asterisk, and affilnum from YAML
\begin{center}
\large
%% removed space here
Lyndon D. Estes\textsuperscript{*} \textsuperscript{1}, 
%% removed space here
Su Ye \textsuperscript{1,2}, 
%% removed space here
Lei Song \textsuperscript{1}, 
%% removed space here
Boka Luo \textsuperscript{1,3}, 
%% removed space here
J. Ronald Eastman \textsuperscript{1,3}, 
%% removed space here
Zhenhua Meng \textsuperscript{1}, 
%% removed space here
Qi Zhang \textsuperscript{1}, 
%% removed space here
Dennis McRitchie \textsuperscript{4}, 
%% removed space here
Stephanie R. Debats \textsuperscript{4}, 
%% removed space here
Justus Muhando \textsuperscript{5}, 
%% removed space here
Angeline H. Amukoa \textsuperscript{5}, 
%% removed space here
Brian W. Kaloo \textsuperscript{5}, 
%% removed space here
Jackson Makuru \textsuperscript{5}, 
%% removed space here
Ben K. Mbatia \textsuperscript{5}, 
%% removed space here
Isaac M. Muasa \textsuperscript{5}, 
%% removed space here
Julius Mucha \textsuperscript{5}, 
%% removed space here
Adelide M. Mugami \textsuperscript{5}, 
%% removed space here
Judith M. Mugami \textsuperscript{5}, 
%% removed space here
Francis W. Muinde \textsuperscript{5}, 
%% removed space here
Fredrick M. Mwawaza \textsuperscript{5}, 
%% removed space here
Jeff Ochieng \textsuperscript{5}, 
%% removed space here
Charles J. Oduol \textsuperscript{5}, 
%% removed space here
Purent Oduor \textsuperscript{5}, 
%% removed space here
Thuo Wanjiku \textsuperscript{5}, 
%% removed space here
Joseph G. Wanyoike \textsuperscript{5}, 
%% removed space here
Ryan B. Avery \textsuperscript{6}, 
%% removed space here
Kelly K. Caylor \textsuperscript{6,7,8}, 
\end{center}
%%Adds affiliations from YAML
\begin{justify}
% \footnotesize \emph{
% \\*
\footnotesize\textsuperscript{1}Graduate School of Geography, Clark
University, Worcester, MA, USA\\\\*
% \\*
\footnotesize\textsuperscript{2}Department of Natural Resources and the
Environment, University of Connecticut, Storrs, CT, USA\\\\*
% \\*
\footnotesize\textsuperscript{3}Clark Labs, Clark University, Worcester,
MA, USA\\\\*
% \\*
\footnotesize\textsuperscript{4}Independent contributor\\\\*
% \\*
\footnotesize\textsuperscript{5}SpatialCollective, Nairobi, Kenya\\\\*
% \\*
\footnotesize\textsuperscript{6}Department of Geography, University of
California Santa Barbara, Santa Barbara, CA, USA\\\\*
% \\*
\footnotesize\textsuperscript{7}Earth Research Institute, University of
California Santa Barbara, Santa Barbara, CA, USA\\\\*
% \\*
\footnotesize\textsuperscript{8}Bren School of Environmental Science and
Management, University of California Santa Barbara, Santa Barbara, CA,
USA\\\\*
% }
%%Adds corresponding author email(s) from YAML
\newcounter{num}
\setcounter{num}{1}
\\[0.1cm]
\footnotesize \emph{
\ifnum\value{num}=1%
\textsuperscript{*} corresponding author:
\fi
\href{mailto:lestes@clarku.edu}{\nolinkurl{lestes@clarku.edu}}
\stepcounter{num}
}

%\begin{center}
This paper is a non-peer reviewed preprint submitted to EarthArXiv. It
will be periodically updated as it is revised, including information on
its peer review status.
%\end{center}

\end{justify}
%%Adds date from YAML
\normalsize

\begin{abstract}
Understanding agricultural change requires reliable, frequently updated
maps that describe the characteristics of croplands. Such data are often
unavailable for regions dominated by smallholder agricultural systems,
which are particularly challenging for remote sensing. To overcome these
challenges, we designed a system to minimize several sources of error
that arise when mapping smallholder croplands. To overcome errors caused
by mismatches between image resolution and cropland scales, as well as
persistent cloud cover, the system converts daily, 3.7 m PlanetScope
imagery into two seasonal composites within a single agricultural year.
To reduce errors that occur when training classifiers, we built a
labelling platform that rigorously assesses label accuracy, and creates
more accurate consensus labels that train a Random Forests model. The
labelling platform and model interact within an active learning process
that boosts the accuracy of the resulting cropland probability map,
which is used in a segmentation process to delineate individual field
boundaries. We applied this system to map Ghana's croplands for the year
2018. We divided Ghana into 16 mapping regions (12,160-23,535 km\(^2\)),
training separate models for each using a total of 6,299 labels, plus
1,600 for validation. Using an independent map reference sample
(n=1,207), we found that overall accuracies of the resulting cropland
probability and field boundary maps were 88\% and 86.7\%, respectively,
with User's accuracies for the cropland class of 61.2\% and 78.9\%, and
Producer's accuracies of 67.3\% and 58.2\%. Croplands covered
16.1-23.2\% of the mapped area, comprising 1,131,146 total fields with
an average size of 3.92 ha. Estimates based on the map reference sample
indicate the cropland percentage is 17.1\% (15.4-18.9\%) or 17.6\%
(15.6-19.6\%), depending on the map used to estimate the standard error.
Using the labellers' digitized field boundaries to estimate biases in
field boundary statistics, we calculated an adjusted mean field size of
1.73 ha and total field count of 1,662,281. Although the cropland class
contained substantial errors, the system was effective in mitigating
error and quantifying resulting performance gains. By minimizing
training errors, consensus labelling improved the model's F1 scores by
up to 25\%, while 3 iterations of active learning increased the F1 score
by 9.1\%, on average, which was 2.3\% higher than training models with
randomly selected labels. Map accuracy can be improved by replacing
Random Forests with a convolutional neural network. These results
demonstrate a readily adapted, transferrable framework for developing
high resolution, annual, nation-scale maps that provide important
details about smallholder-dominated croplands.
\end{abstract}
\end{singlespace}

\bleft

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Amidst all the challenges posed by global change, a particular concern
is how agricultural systems will adapt to meet humanity's growing food
demands, and the impacts that transforming and expanding food systems
will have on societies, economies, and the environment (Searchinger et
al. 2019). Significant efforts are being made to address the various
aspects of this challenge, including work on diagnosing and closing
yield gaps (Lobell et al. 2009, e.g. Licker et al. 2010, Mueller et al.
2012), expanding and commercializing production (Morris and Byerlee
2009), and to understand (Rulli and D'Odorico 2014, Kehoe et al. 2017,
Davis et al. 2020) and mitigate (Estes et al. 2016b) agriculture's
ecological impacts. Answering many of the questions these efforts seek
to address depends on reliable data that describes the location and
characteristics of cropland (Fritz et al. 2015), and how these are
changing over time. Unfortunately, the data that do exist are in many
places inaccurate. Existing estimates of how much global cropland there
is tend to vary widely, and they often disagree about where cropland is
located (e.g. Fritz et al. 2011, 2013). Such errors can propagate in
subsequent analyses that rely on cropland data as inputs, resulting in
potentially misleading answers (Estes et al. 2018). Beyond cropland
distributions, few data are available on key cropland characteristics
such as field size, an important variable needed to estimate yield and
other key food security variables (Carletto et al. 2015), and as an
indicator of farm size (Levin 2006, Samberg et al. 2016), a critical
component of rural livelihoods given increasing population densities and
longstanding debates about the relationship between farm size and
productivity (Feder 1985, Carletto et al. 2013, Desiere and Jolliffe
2018).

These informational inadequacies are due to the fact that cropland data
in much of the world are derived from remotely sensed landcover maps,
which can be notoriously high in error, particularly over regions such
as Africa (Fritz et al. 2010, Estes et al. 2018), where agricultural
changes will be largest and the need for accurate baseline data is thus
greatest (Searchinger et al. 2015, Estes et al. 2016b, Bullock et al.
2021). Cropland mapping over Africa is difficult for several reasons.
The primary reason relates to the characteristics of the continent's
smallholder-dominated croplands, where half of all fields are smaller
than 1 ha (Lesiv et al. 2019). This size is small relative to the 30-250
m resolution of the sensors typically used in many landcover mapping
efforts (e.g. Chen et al. 2015, Sulla-Menashe et al. 2019), which
results in errors due to mixed pixels and aspects of the modifiable area
unit problem (Openshaw and Taylor 1979). In the latter case, the pixel's
shape may be poorly matched to that of cropland, and is too coarse to
aggregate to approximate that shape at the characteristic scales of crop
fields (Dark and Bram 2007, Estes et al. 2018). On top of the matter of
scale is 1) high intra-class variability of the cropland class,
compounded by the fact that these particular croplands can be heavily
intergraded with surrounding vegetation (Debats et al. 2016, Estes et
al. 2016a), and 2) the substantial temporal variability within
croplands, both within and between seasons. These latter two aspects
pose challenges for the classification algorithms that are applied to
the imagery.

These problems arising from cropland characteristics are increasingly
being addressed due to technological advances. Recent advances in
satellite technology have increased the coverage of high (\textless5 m)
or near-high (10 m) resolution imagery with weekly to near-daily return
intervals (Drusch et al. 2012, McCabe et al. 2017). This high spatial
\emph{and} temporal resolution addresses the sensor-field scale
mismatch, and more effectively captures the intra-seasonal dynamics of
cropland, which helps classifiers distinguish cropland from surrounding
cover types (Debats et al. 2016, Defourny et al. 2019). On top of this,
advances in cloud computing (Gorelick et al. 2017), the opening of image
archives (Wulder et al. 2016), and next generation machine learning
approaches (Zhu et al. 2017, Maxwell et al. 2018) are placing large
volumes of these moderate to near-high resolution imagery together with
the computational and algorithmic resources necessary to classify them
at scale. These capabilities are aleady being used to create a new
generation of higher resolution (10-30 m) cropland and landcover maps
for Africa and other regions (Xiong et al. 2017, Lesiv et al. 2017, ESA
n.d.).

Despite these advances, the highest resolution (\textless5 m) image
sources are still not used to map cropland over very large extents,
presumably because they are commercial and relatively high cost to
acquire, in addition to the greater computational demands. As such, map
accuracy can still be a challenge, particular for User's accuracy, which
ranges between 46 and 76\% for the cropland class (Xiong et al. 2017,
e.g. Lesiv et al. 2017).

Accuracy may also suffer due to error-inducing factors that are becoming
somewhat more pronounced as a consequence of these technology advances,
particularly with respect to algorithms. Advances in machine learning
are helping to greatly improve classification skill, but these
algorithms generally require large, high-quality training datasets
(Maxwell et al. 2018, Ma et al. 2019, Elmes et al. 2020). To satisfy
this need for more training (and reference) samples, map-makers
increasingly rely on visual interpretation of high resolution satellite
or aerial imagery to collect training (or validation) samples (Chen et
al. 2015, e.g. Xiong et al. 2017, Stehman and Foody 2019). Several
web-based platforms have been developed to facilitate such efforts,
which provide convenient and highly scalable tools for training data
collection (Fritz et al. 2012, Estes et al. 2016a, e.g. Bey et al.
2016). Visually interpreted training labels present two particular
problems. The first is that such labels have inevitable interpretation
errors that can vary substantially according to the skill of the
interpreter (Estes et al. 2016a, Waldner et al. 2019). These errors are
typically not accounted for in reported accuracy metrics, despite the
fact that they can introduce substantial error into the resulting maps
and subsequent analyses (Estes et al. 2018, Elmes et al. 2020). The
second problem is that visual interpretation depends on high resolution
imagery (\textless5 m), as lower resolutions make it difficult for a
human analyst to discern cropland. Typically the only practical source
for such imagery are ``virtual globe'' basemaps provided by Microsoft
and Google, which are composed of mosaics of various high resolution
satellite and aerial images that typically span 3-5 years of time within
a single country (Lesiv et al. 2018). This within-mosaic temporal
variation can set up a temporal mismatch between the imagery being
interpreted and the imagery being classified, which is usually from a
different source (e.g.~Landsat, Sentinel; Xiong et al. (2017)). If a
land change occurs in the interval between the two image sets (e.g.~a
new field was created), this can introduce error into the training data
that is then passed on to the classifier. This source of error may be
elevated in smallholder-dominated systems in the tropics, where swidden
practices are common (Van Vliet et al. 2013), or in rapidly developing
agricultural frontiers (Zeng et al. 2018).

Improving the accuracy of cropland maps over smallholder-dominated
systems thus requires an approach that meets three requirements. First,
it should be based on high spatial and temporal resolution imagery, to
be able to capture the fine grain and temporal variability of
smallholders' fields. Second, an algorithm with suitable skill for
classifying these images must be selected, and combined with the
computational resources needed to process large imagery volumes. Third,
a method for collecting large volumes of high quality training and
validation data based on image interpretation is essential. This method
should quantify and minimize the errors associated with image
interpretation. It should also ensure that labels are collected either
from the same imagery that is being classified, or from contemporanous
imagery, in order to reduce errors introduced by land change processes.

We describe here a cropland mapping system that follows these
requirements, with an emphasis on delineating field boundaries. The
first requirement is enabled by the recent availability of CubeSat data
that provides 3-4 m resolution imagery over large areas at near daily
intervals (McCabe et al. 2017). Although these data, currently collected
by Planet's CubeSat fleet, are of lower spectral depth and quality than
Landsat, Sentinel, or Worldview imagery, they enable country to
continental scale image mosaics to be created for multiple periods
during the crop growing calendars, and capturing the intra-annual
variability can be more important for classifying cropland than spectral
depth (Debats et al. 2016). This daily revisit capacity is also
important for developing seasonal composites in cloudy regions, where
satellites with longer return intervals may fail. Lastly, although this
imagery being up to \textasciitilde16 times coarser than much of Bing or
Google imagery, it is sufficiently resolved for humans to discern most
fields (Fourie 2009, e.g.~see Estes et al. 2018). This allows labels to
be generated on the same imagery processed by the classifier, thereby
addressing one of the two needs related to training data (requirement
3).

The second requirement is addressed by a computer vision/machine
learning classifer that is effective for classifying smallholder
croplands (Debats et al. 2016), re-engineered to run on high
performance, cloud-based computing clusters with a simplified feature
set, and following recommended practices for controlling for and
measuring the errors that occur when training machine learning models
for remote sensing applications (Elmes et al. 2020). The classifier is
tightly coupled to a front-end platform for collecting label data, which
includes rigorous accuracy assessment protocols and a novel approach for
merging multiple maps into consensus labels, thereby minimizing image
interpretation error (Estes et al. 2016a, Elmes et al. 2020). The
training and machine learning components are combined within an active
learning framework, wherein the machine learning process assesses
classification uncertainty in unlabelled areas after a training step,
and selects sites from areas of highest uncertainty for additional
labelling (Cohn et al. 1994, Tuia et al. 2011). Our framework automates
this interactive approach to label selection, which is effective in
boosting the performance of classification models while reducing the
overall number of training samples required to achieve a given level of
performance (Debats et al. 2017, e.g. Hamrouni et al. 2021). Finally, an
unsupervised segmentation step is applied to the imagery and merged with
the pixel-wise classifications from the machine learning process,
resulting in a vectorized field boundary map that provides important
information on field geometry.

We demonstrate this approach to map cropland in Ghana, a country where
smallholder farming predominates, and which has a broad mix of climates
and agricultural systems, including large areas where shifting
agriculture is practiced (Samberg et al. 2016, Kansanga et al. 2019).

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{system-overview}{%
\subsection{System overview}\label{system-overview}}

The mapping system design has four primary components focused on 1)
image acquisition and compositing, 2) training and validation data
collection, 3) image classification, and 4) segmentation. The first
component is applied as a once-off step to generate an image catalog
covering the mapping geography. The second component provides tools for
labelling imagery, and controls an interactive model training and
classification (component 3) pipeline that produces a map of cropland
probabilities for each image pixel (Figure \ref{fig:systemoverview}).
The final segmentation step (component 4) is then initiated and applied
to both the input image catalog and the posterior probability maps,
resulting in vectorized field boundaries. Each system component
comprises an individual software module designed to be implemented on
cloud computing architecture, and available on a GitHub repository (see
data and software availability section for details).

\begin{figure}

{\centering \includegraphics[width=0.5\linewidth]{figures/figure1}

}

\caption{An overview of the primary system components, the data stores that hold the inputs and outputs from each component, and the direction of connections between them. The dashed line indicates iterative interactions, while solid lines indicate one-time or irregular connections.}\label{fig:systemoverview}
\end{figure}

\hypertarget{system-components}{%
\subsection{System components}\label{system-components}}

\hypertarget{image-compositing}{%
\subsubsection{Image compositing}\label{image-compositing}}

The image processing components of our system were designed to work with
PlanetScope Analytic surface reflectance data (PlanetTeam 2018).
PlanetScope provides three visual (red, green, blue) and a near-infrared
band at 3-4 m resolution at nominal daily frequency. Although these
images are already pre-preprocessed and corrected for atmospheric
effects, there are residual errors from inter-sensor differences and the
radiometric normalization process (Houborg and McCabe 2018), variation
in the orientation of scene footprints, as well as a high frequency of
cloud cover over the study region (Wilson and Jetz 2016). To correct for
these factors, we developed a procedure for creating temporal composites
representing the primary growing and non-growing seasons within a single
year.

PlanetScope imagery is accessed via the Planet API (PlanetTeam 2018),
and an initial order is placed for all imagery falling within the
mapping geography and the date ranges for the two compositing periods.
The imagery is collected and transferred directly to a cloud storage
platform (Amazon Web Services {[}AWS{]} S3).

Individual images are then transformed into analysis ready data (ARD)
(Dwyer et al. 2018), by subsetting each downloaded image into 0.05
degree tiles, regardless of cloud cover. Tiles are organized within a
larger 1 degree resolution grid that covers the entire continent, which
defines the minimum mapping area of interest (AOI; Figure
\ref{fig:aois}).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figures/figure2}

}

\caption{The reference system used in the mapping platform, including A) numbered areas of interest (AOIs) that define the minimum mapping geography (solid black lines; dotted lines indicate boundaries of 1 degree grid), B) the 0.05 degree tile used for compositing imagery, and C) the 0.005 degree resolution reference grid used for collecting training data and distributed computing.}\label{fig:aois}
\end{figure}

The temporal compositing process is applied to the tiled daily images
for the time period of interest, which in this case was one of two
multi-month seasons, the primary growing and dry seasons for a single
agricultural year. Imagery from two seasons helps to improve the
performance of cropland classifiers (Debats et al. 2016), while having
the seasons in the same year helps to minimize differences caused by
land change. For each pixel in each image in each ARD temporal stack for
a given season, two weights are calculated:

\begin{equation} \label{eq:cloud}
\mathrm{W1_t} = \frac{1}{\mathrm{blue_t}^2}
\end{equation}

\begin{equation} \label{eq:shadow}  
\mathrm{W2_t} =\begin{cases}
    \frac{1}{\mathrm{NIR_t}^4}, & \text{if $\mathrm{NIR_t}$ < median\{$\mathrm{NIR_{t1}}$, $\mathrm{NIR_{t2}}$, ..., $\mathrm{NIR_{ti}}$\}}.\\
    1, & \text{otherwise}.
  \end{cases}
\end{equation}

Where \emph{t} is a particular date in near-daily time series of
PlanetScope images, which begins at date 1 for the given compositing
period and ends on date \emph{i}, \emph{blue} is the blue band, and
\emph{NIR} the near infrared band. Equation \ref{eq:cloud} assigns lower
weights to hazy and clouded pixels as the blue band is sensitive to haze
and cloud pixels (Zhang et al. 2002), while Equation \ref{eq:shadow}
assigns low weights to pixels in cloud shadow considering the
significant darkening effect of the cloud shadows in the Near Infrared
band (Zhu and Woodcock 2012, Qiu et al. 2020)

Once these two weights are calculated, the final composited pixel value
for each of the four PlanetScope bands is:

\begin{equation}
\mathrm{\bar{B} = \frac{\sum_{t=1}^{T}B_t * W1_t * W2_t}{\sum_{t=1}^{T}W1_t * W2_t}}
\end{equation}

Which is a weighted mean for each pixel for each band \emph{B} for the
particular compositing period. The composited tiles were then added to
the S3 store (Figure \ref{fig:systemoverview}), where they are stored as
cloud-optimized geotiffs, and a ``slippy
map\footnote{https://wiki.openstreetmap.org/wiki/Slippy\_Map}''
rendering is created for each composite using Raster Foundry (Azavea
2020). The web-rendered imagery is presented within the training data
platform (next section).

\hypertarget{labelling-platform}{%
\subsubsection{Labelling platform}\label{labelling-platform}}

Training and reference data are collected by a custom labelling
platform, which was originally designed for AWS's Mechanical Turk job
marketplace (Estes et al. 2016a). The basic structure of the system
remains the same, but we converted it into a standalone platform that
allows us to enroll and pay people directly for their labelling, and is
designed to control and supervise the machine learning process. The
platform runs on a Linux virtual machine hosted on an AWS EC2 instance
and is comprised of a database (PostGIS/Postgres), a mapping interface
(OpenLayers 3), an image server (Raster Foundry), and a set of utilities
for managing, assessing, and converting digitization work into
rasterized labels for training a machine learning algorithm. Each
instance of the platform focuses on a specific AOI (Figure
\ref{fig:aois}A)

The following sections provide an overview of the labelling platform's
architecture.

\hypertarget{mapping-workflow}{%
\paragraph{Mapping workflow}\label{mapping-workflow}}

\hypertarget{selecting-training-and-reference-sites}{%
\subparagraph{Selecting training and reference
sites}\label{selecting-training-and-reference-sites}}

The labelling process begins with the random selection of a subset
(e.g.~500) of cells from a 0.005 degree grid, with the selection itself
potentially split into a training and validation sample, according to
predetermined proportions. The grid, which is nested within the tiling
and larger 1 degree grids (Figure \ref{fig:aois}C) defines the spatial
unit for a labelling job. The selected cells are placed into a queue
within the platform's database, and then converted into a mapping
\emph{task} that has a specified number of \emph{assignments}
(boundaries drawn by an individual labeller) that must be completed
before the task is complete.

\hypertarget{mapping-assignments}{%
\subparagraph{Mapping assignments}\label{mapping-assignments}}

Labellers registered in the system log in to the mapping platform (built
with Flask) and navigate to the OpenLayers-based field mapping interface
(Figure \ref{fig:labeller}), where they are presented with a white
target box representing the randomly selected grid cell, a set of
digitizing tools, and different image backdrops, including true and
false color renderings of the growing season and off-growing season
PlanetScope composites, and several virtual globe basemaps. Following a
set of pre-defined digitizing rules (see SI), the labeller uses the
polygon drawing tool to digitize the boundaries of all crop fields
intersecting the target grid cell that are visible within the
PlanetScope overlays. To aid with interpretation, the labeller can
toggle between the PlanetScope renderings and the basemaps to help form
a judgement about what constitutes a field. The labeller assigns each
polygon a class category (e.g.~annual cropland), and upon completing all
fields submits the assignment to the database. In cases where the target
grid cell does not contain any fields, the labeller simply submits the
assignment to mark it complete. The labeller is then directed to the
next available assignment from a different labelling task.

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{figures/figure3}

}

\caption{An overview of the labelling platform's interface}\label{fig:labeller}
\end{figure}

\hypertarget{processing-completed-assignments}{%
\subparagraph{Processing completed
assignments}\label{processing-completed-assignments}}

All submitted polygons are cleaned to fix topological irregularities
that arose during digitization (see supporting information {[}SI{]}) and
stored in a PostGIS table. Each completed assignment represents one of
two types of tasks: 1) accuracy assessment, or 2) model training or
validation. For the former type, an accuracy assessment routine is
invoked that executes a series of comparisons between the labeller's
results and a training reference dataset, resulting in a assignment
score:

\begin{equation} \label{eq:qaqc}
\mathrm{score_i}=\beta_0\mathrm{I}+\beta_1\mathrm{O}+\beta_2\mathrm{F}+\beta_3\mathrm{E}+\beta_4\mathrm{C}
\end{equation}

Where \emph{i} indicates the particular assignment, and \(\beta_{0-4}\)
represent varying weights that sum to 1. \emph{I} refers to ``inside the
box'' accuracy, \emph{O} is the accuracy of those portions of the
labeller's polygons extending beyond the target grid boundaries,
\emph{F} is fragmentation accuracy, a measure of how many individual
polygons the labeller delineated relative to the reference, \emph{E}
measures how closely each polygon's boundary matched its corresponding
reference polygon boundary, and \emph{C} assesses the accuracy of the
labeller's thematic labels (see SI for individual formulae). Equation
\ref{eq:qaqc} is an extension of the approach described by Estes et
al.~(2016).

Over time, labellers are assessed multiple times across a range of
accuracy tasks, which are selected to represent the variability of the
agricultural system being mapped. Each labeller's score history is
averaged to provide an overall accuracy measure, and this information is
used for creating labels, the second task.

If the labeller's completed assignment was a training/validation task,
their maps remain stored in the database until the task's outstanding
assignments are completed by other labellers. Once complete, another
routine is invoked, which combines the task's completed assignments into
a single consensus label using a Bayesian merging approach:

\begin{equation}
P(\theta|\mathrm{D})=\sum_{i=1}^{n}\mathrm{P}(\mathrm{W_i}|\mathrm{D})\mathrm{P}(\mathrm{\theta}|\mathrm{D}, \mathrm{W_i})
\end{equation}

Where \(\theta\) represents the true cover type of a pixel (field or not
field), \emph{D} is the label assigned to that pixel by a labeller, and
\(W_i\) is an individual labeller. P(\(\theta\)\textbar D) is therefore
the probability that the actual cover type is what the labellers who
mapped it says it is, while P(W\(_i\)\textbar D) is an individual
labeller's average score over all the accuracy assessment assignments
they have completed, and P(W\(\theta\)\textbar D, \emph{W\(_i\)}) is the
labeller's label for that pixel. This approach therefore uses the
overall accuracy of each labeller to weight their labels when combined
with those made by other labellers' for the same pixel (see SI for
further details). As a further measure of confidence in the final
consensus label, its average Bayesian Risk can be calculated (see SI).
This measure ranges between 0 and 1, with 0 indicating full agreement
between labellers for all pixels (n = 40000) in the label, and 1
indicating complete disagreement.

\hypertarget{classification-pipeline}{%
\subsubsection{Classification pipeline}\label{classification-pipeline}}

Upon completion of a batch of labels, the platform automatically
launches an ephemeral Elastic Map
Reduce\footnote{https://docs.aws.amazon.com/emr/latest/APIReference/emr-api.pdf}
cluster consisting of tens of instances, depending on the size of the
AOI.

\hypertarget{feature-extraction}{%
\paragraph{Feature extraction}\label{feature-extraction}}

The first step is the extraction of additional features from each
seasonal image composite. Previous work showed that a large number of
simple features that summarize the statistical properties of reflectance
and vegetation indices in local neighborhoods are highly effective for
classifying smallholder croplands (Debats et al. 2016). We followed this
logic in this study, but were constrained to use a smaller feature space
because the storage and memory requirements for our mapping geographies
in this case were several orders of magnitude larger. For this
implemention, we thus extract a set of 16 features, which are the mean
and standard deviations calculated within an 11X11 and 5X5 moving
window, respectively (initial tests revealed these two window sizes to
be most effective), resulting in 24 overall features, including the
original bands (Table 1).

\begin{center}Table 1. List of image features.\end{center}

\begin{longtable}[]{@{}lll@{}}
\toprule
Feature & Window Size & N Features \\
\midrule
\endhead
RGB-NIR & 1X1 & 8 \\
Mean & 11X11 & 8 \\
Standard deviation & 5X5 & 8 \\
\bottomrule
\end{longtable}

Feature extraction and the conversion of image features is handled by a
combination of
\texttt{GeoTrellis}\footnote{https://github.com/locationtech/geotrellis},
\texttt{rasterio}\footnote{https://rasterio.readthedocs.io/en/latest/},
and \texttt{RasterFrames}\footnote{https://rasterframes.io/}. These
collectively extract subsets of imagery from the PlanetScope temporal
composites, derive the features, and convert these into Apache Spark
DataFrames. Features are extracted on the fly for each cell in the
training and validation sets, a functionality enabled by storing the
image composites as Cloud-optimized
Geotiffs\footnote{https://www.cogeo.org/} (COGs).

\hypertarget{classification}{%
\paragraph{Classification}\label{classification}}

Once the features from the training sites are extracted into
\texttt{RasterFrames}, these are combined with their corresponding
labels and passed to the machine learning classifier, a
\texttt{SparkMLlib} implementation of Random Forests (Breiman 2001). For
this study, the model was trained with a balanced sample and a tree
depth of 15 and total tree number of 60, which initial testing showed to
provide a reasonable balance between computational time and performance.

After fitting, the model is applied to the features of the model
validation set, and a set of performance metrics is calculated,
including binary accuracy, the F1 score (the geometric mean of precision
and recall), and the area under the curve of the Receiver Operating
Characteristic (Pontius and Si 2014).

\hypertarget{the-active-learning-loop}{%
\subsubsection{The active learning
loop}\label{the-active-learning-loop}}

After fitting and model evaluation a second prediction is undertaken to
enable active learning. The feature extraction process is repeated for
the rest of the mapping geography that falls outside of the training and
validation sample, but applied to a subset of randomly drawn pixels from
each cell in order to reduce computational demand. The fitted model is
applied to predict the cropland probability for these selected pixels,
and an uncertainty criterion (Debats et al. 2017) is calculated for each
grid cell:

\begin{equation}
\mathrm{Q_I = \sum_{I(x, y) \epsilon I} (p(x, y) - 0.5)^2}
\end{equation}

Where Q is the uncertainty for grid cell I, calculated from the
predicted probability \emph{p} of the randomly selected subset of pixels
(x, y) drawn from it. Pixels with predicted probabilities closer to 0.5
are least certain as to their classification, thus images with the
lowest values of \(Q\) represent sites posing the most difficulty for
the classifier.

After scoring with the uncertainty criterion, the top \emph{N} most
uncertain grid cells are selected and sent back to the labelling
platform, which are then digitized by the labellers. The resulting
consensus labels from the actively selected sample are added to the
initial randomly selected sample, and a new cluster is launched. The
model is retrained, assesses uncertainty across the remaining unlabelled
sites, and selects the next most uncertain sites for labelling. This
loop repeats until model performance gains against the validation set
show diminishing returns.

\hypertarget{segmentation}{%
\subsubsection{Segmentation}\label{segmentation}}

After the final iteration, the segmentation algorithm is invoked, which
entails several steps. In the first step, the meanshift algorithm
(Yizong Cheng 1995) is applied to the original bands of the dry season
composite. A Sobel filter is then applied to the green, red, and
near-infrared mean-shifted bands and the probability map, and a combined
edge image is computed using the sum of these four edge images for the
dry season only. A compact watershed algorithm (Neubert and Protzel
2014) is then run on the weighted edge image, with a high level of
segmentation specified. In this case, we specified \textbf{6400}
segments per tile.

Third, a region adjacency graph is constructed for each image tile, in
which each node represents all pixels within each polygon created in the
previous step. The edge between two adjacent regions (polygons) is
calculated as the norm of the difference between the means of normalized
colors of all bands. Hierarchical merging is then applied, in which the
most similar pairs of adjacent nodes are merged until there are no edges
remaining below a predetermined threshold of 0.05.

In the fourth step, the merged polygons are overlaid with the posterior
probabilities resulting from the final active learning loop, and
polygons in which the average posterior probability is greater than a
predetermined threshold (here 0.5, but could vary locally) are retained
as field polygons.

In the final step, the retained polygons are refined by removing holes
and smoothing their boundaries using the Visvalingam algorithm
(Visvalingam and Whyatt 1993). Neighboring polygons that overlap along
tile boundaries are then merged.

To assess the accuracy of the final segmented boundaries, we used a
two-step approach. First, we assessed the overall thematic accuracy of
the resulting classification against our map reference data. Second, to
assess the quality of the segmentation, we compared the mean area and
relative frequencies of the segmented polygons within different size
classes against the same metrics derived from the digitized fields of
the most accurate worker to create the given map. We selected this
relatively simple procedure, as opposed to more complex measures of
object accuracy (Ye et al. 2018), because, on the one hand, both the
automated segmentation algorithm and labellers are cueing in on the same
features--abrupt, physically detectable breaks within the imagery. On
the other hand, no matter how well the intepreted/segmented boundaries
align with the boundaries of fields in the imagery, it is logistically
difficult to evaluate performance against real-world boundaries as the
spectral distinction of field boundaries will vary across different crop
types and land use arrangements.

\hypertarget{applying-the-system-to-map-ghana}{%
\subsection{Applying the system to map
Ghana}\label{applying-the-system-to-map-ghana}}

We applied the system to map Ghana's croplands, excluding areas
primarily cultivated with tree crops. Ghana has several distinct
agricultural regions, ranging from the primarily grain and vegetable
crop producing regions in the northern savannas to tree crop-dominated
system in the forested southwest, where cocoa and oil palm are among the
dominant crops. For these latter regions, we did not attempt to classify
tree crops, and instead mapped clearings that potentially contain field
crops or newly felled or recently replanted tree crops. We made this
decision because PlanetScope's resolution is not high enough for
labellers to distinguish many tree crops from surrounding forest, and
the boundaries of many tree crops (e.g.~cocoa) are often not visible.

To create the cropland maps, we divided the country into 16 distinct
AOIs, which were developed by grouping together each one degree cell
fully contained within Ghana with the tiles belonging to any adjacent
degree cell that overlapped neighboring countries (Figure
\ref{fig:aois}A). The exception was AOI 16, which consisted of the four
degree cells intersecting Ghana's southern coast. The resulting AOIs
ranged from 12,160 to 23,535 km\(^2\) in extent (average = 15,457
km\(^2\)), A separate active learning and segmentation process was run
for each of these AOIs.

To collect the initial randomized samples for model training, we grouped
the AOIs into three clusters: a northern cluster comprising the 6
northernmost AOIs (Cluster 1), a central to southeastern cluster
(Cluster 2) consisting of the 3 middle (AOIs 7-9) and 2 southeastern
AOIs (12 and 15), and a southwestern cluster (Cluster 3) made up of the
forest zone AOIs (10, 11, 13, 14, 16). Within each cluster, we randomly
selected and labelled 500 grid cells, which provided relatively large
initial training samples for these agro-ecologically similar regions,
while helping to minimize the overall amount of labelling effort. In
addition to these samples, we randomly selected and labelled 100 grid
cells within each AOI to provide a validation sample.

After collecting the initial training and validation samples, we trained
a starter model for each cluster and applied it to each of the block's
AOI. For each iteration, 100 samples were actively selected within each
AOI, and added to the training pool.

During the collection of training and validation samples, labellers were
tasked to only digitize active or recently active crop fields, avoiding
tree crops, and fallow or potentially abandoned fields (see SI for the
digitizing rules).

To evaluate the performance of the system, we performed several analyses
described in sections 2.3.1-4.

\hypertarget{image-quality}{%
\subsubsection{Image quality}\label{image-quality}}

We evaluated the overall quality of the resulting seasonal image
composites by assessing a random selection of 50 tiles. We graded both
seasonal composites for each tile using a four category quality score,
which evaluated the degree of 1) residual cloud and 2) cloud shadow, 3)
the number of visible scene boundary artifacts, and 4) the proportion of
the image that had its resolution degraded below the typical 3-4 m
PlanetScope resolution (e.g.~because of between-date image
mis-registrations). Each category was qualitatively ranked from 0-3,
with 0 being the lowest quality, and 3 the highest (see SI for complete
protocol), making the highest possible score 12. We rescaled scores to
fall between 0 and 1.

\hypertarget{model-gains-per-iteration}{%
\subsubsection{Model gains per
iteration}\label{model-gains-per-iteration}}

To assess the gain in model performance due to active learning, we
measured the change in accuracy, F1, and AUC (see 2.2.3.2) between each
iteration and between the first and last iterations for each AOI.

To evaluate whether active learning improved model performance relative
to a purely random approach to selecting new training sites, we ran
additional tests within a subset of AOIs (1, 8, and 15). We first
randomly selected and labelled 300-400 sites in each AOI. We then
progressively added 100 of the randomly selected samples to the relevant
training pool and retrained the model, repeating the process so that the
number of iterations and samples matched those from the active learning
process. We then compared the difference in accuracy, AUC, and F1
between the randomly trained models and those trained with active
learning (Debats et al. 2017).

\hypertarget{accounting-for-label-error}{%
\subsubsection{Accounting for label
error}\label{accounting-for-label-error}}

To quantify the potential impact of label error on classification
results, we evaluated the performance differences between models trained
with three different sets of labels: 1) those from the lowest scoring
labeller to map each training site, 2) those from the highest scoring
labeller, and 3) the consensus labels. This assessment follows
recommended Tier 1 (i.e.~best practice) standards to account for
training data errors (Elmes et al. 2020).

\hypertarget{accuracy-assessment}{%
\subsubsection{Accuracy assessment}\label{accuracy-assessment}}

The model performance assessments described above (2.3.2-3) were not
fully independent because they used the same validation sites over
multiple iterations (Elmes et al. 2020). To independently assess the
accuracy of our final map products, we followed recommended guidelines
(Stehman and Foody 2019) to create a separate map reference sample. We
used a stratified design, randomly assigning square polygons of
\textasciitilde0.1 ha extent into cropland and non-cropland strata,
developed the map of segmented field boundaries. Four classes were used
for the map reference sample: cropland; non-cropland; unsure but likely
cropland; unsure but likely non-cropland. The latter two classes were
used to provide insight into the degree of uncertainty in the map
reference sample. For efficiency, two separate supervisors evaluated
separate portions of the reference sample, but both jointly assessed a
small subset of the sample. We calculated their level of agreement on
this subset to provide an additional assessment of uncertainty in the
map reference sample (Stehman and Foody 2019). The SI contains further
details on the design and collection of the map reference sample.

The map reference polygons were then intersected with both the
probability images and the segmented field boundaries, and confusion
matrixes between the map reference labels and the extracted map classes
were constructed to assess the categorical accuracy of each map product.
We calculated the overall accuracy for each map, as well as the
class-wise User's and Producer's accuracy, as well as the 95\%
confidence intervals for each accuracy measure (Olofsson et al. 2013,
2014, Stehman and Foody 2019).

To assess the accuracy of the segmented field boundaries, we compared
the size class distributions of the segmented field boundaries against
those of the workers' digitized polygons at map validation sites. We
chose this approach because of existing uncertainties in polygon-based
accuracy assessment methods (Ye et al. 2018), and because the map's
ability to represent field sizes was of greatest interest. To undertake
this comparison, we selected the field polygons from the most accurate
labeller to digitize each of the 100 validation sites in each AOI, and
calculated the site-wise average area and number of polygons. We then
calculated the same statistics from the segmented boundaries that
intersected each validation grid. We compared the distributions and
proximity of two measures of central tendency (mean and median)
calculated from the two datasets for each AOI, and across all AOIs.

\hypertarget{assessing-the-characteristics-of-ghanaian-cropland}{%
\subsection{Assessing the characteristics of Ghanaian
cropland}\label{assessing-the-characteristics-of-ghanaian-cropland}}

Using the final mapped results, we calculated the estimated area of
cropland in Ghana, as well as the average size and total number of
fields in the different AOIs. We used the map reference sample to
calculate adjusted area estimates and confidence intervals for each map
class, and used the differences between labellers' polygons and
segmented boundaries at validation sites to calculate bias-adjusted
estimates of mean field sizes and the total number of fields.

\hypertarget{results}{%
\section{Results}\label{results}}

We developed maps of Ghana's cultivated croplands within an area of
247,299 km\(^2\), which included portions of neighboring countries
overlapped by images tiles.

\hypertarget{image-catalog-and-quality}{%
\subsection{Image catalog and quality}\label{image-catalog-and-quality}}

To develop the maps, we first generated an image catalog for the 8,116
tiles covering Ghana. This entailed processing all PlanetScope imagery
intersecting these tiles between May-September, 2018 (the growing
season) and December, 2018 to February, 2019 (the subsequent dry
season). The longer period was necessary for the growing season because
of the frequent cloud cover, which substantially limits the number of
clear scenes for any tile (Figure S3). For the cloudiest regions (AOIs
10, 11, 13, 14, 16) we started the dry season window in November.

An assessment with two observers (see SI for observer details) found
that average quality per growing season composite tile was 0.88, with 70
percent having scores \(\geq\) 0.85, while the average quality of dry
season composites was 0.92 (74 percent \(\geq\) 0.85). Composite quality
in both seasons was highest in the northern half of the country and
lowest in the southwest (Figure \ref{fig:imqual}A), where the
substantially greater cloud cover resulted in a much lower density of
available PlanetScope imagery for each time period (Figure S3).

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{figures/figure4}

}

\caption{The location and quality scores of 100 randomly selected tiles for the growing (A) and off-growing season (B), and the corresponding distributions of the quality scores for each season, respectively (C and D).}\label{fig:imqual}
\end{figure}

\hypertarget{active-learning}{%
\subsection{Active learning}\label{active-learning}}

\hypertarget{training-data-collection}{%
\subsubsection{Training data
collection}\label{training-data-collection}}

After training of models with the initial randomly selected label sets,
the active learning process was run for 3 iterations for 12 of the 16
AOIs, resulting in 800 labels per AOI. AOIs 10 and 14 stopped after one
and two iterations, respectively, as they started with high initial
validation accuracies (\textgreater83\%) and showed little subsequent
improvement. The models for these two AOIs were thus trained with 600 -
700 samples. AOI 15 was run for 4 iterations (900 samples), while AOI 3
underwent a second active learning cycle because the model produced
during the first cycle was inaccurate (see SI). In this second run, 300
initial training sites randomly selected within the AOI were used
(Figure S4A), followed by 2 subsequent active learning iterations,
resulting in a training sample of 500. Labels collected during the
active learning iterations showed distinct patterns in several AOIs,
which often fell along ecotones, such as the boundaries between
agroecozones (see Figure S4A). The total number of unique training and
validation sites across the country were 6,299 and 1,600, respectively.

The distribution of training and validation sample collection effort was
divided across 20 labellers, with a core group of 13 who mapped more
than 1,000 sites each. As each training site was mapped by 4 separate
labellers, 34,014 sets of vector labels were made. Each labeller
digitized an average of 2,001 (see Figure S5A for more details on
labelling effort). Labeller accuracy was scored 9,389 times against 98
unique training reference sites (Figure S4A), with each labeller
assessed an average of 552 times at a rate of 1 training reference site
for every 3.62 training site mapped. The mean of each labeller's average
accuracy score was 0.71 (range 0.6 to 0.85; see Figure S5B for detailed
score distributions).

After each site was mapped by four labellers, consensus labels were
generated. The Bayesian Risk (see SI) of each training and validation
label was calculated as an additional measure of label quality. The
average risk was 0.122 for training labels and 0.127 for validation
labels. Risk was highest in the northen AOIs (AOIs 1-6; Figures S6-7),
falling between 0.157 for training and 0.173 for validation labels
(Figures S6-7), and lowest in the southwestern AOIs (AOIs 10, 11, 13,
14, 16; training risk = 0.079; validation risk = 0.065). Label risk in
the central-southeastern AOIs (AOIs 7-9, 12, 15) was slightly lower
(training = 0.127; validation = 0.136) than in the north. Labeller
experience also appeared to reduce risk, which we observed during a
relabelling of the 500 initial random site in this cluster (see SI); the
mean risk of the updated labels was 0.055, compared to 0.172 for
original labels.

\hypertarget{performance-gains-during-active-learning}{%
\subsubsection{Performance gains during active
learning}\label{performance-gains-during-active-learning}}

Model performance was calculated for each iteration within each AOI. The
average accuracy, AUC, and F1 at iteration 0 were 0.786, 0.809, and
0.464, respectively, increasing to 0.825, 0.818, and 0.507 by iteration
3 (Figure \ref{fig:alperformance}). These differences represent
respective gains of 4.9, 1.1, and 9.1 percent for the three metrics. The
largest gains for each metric occurred on iteration 1, averaging 2.9, 1,
and 3.8 percent for accuracy, AUC, and F1, while the lowest gains were
realized on iteration 3, with accuracy, F1, and AUC respectively
increasing by just 1.2\%, 0.9\%, and 0.3\%. The scores achieved on the
final iteration varied substantially across AOIs and metrics. Accuracy
ranged between 0.725 (AOI 15) and 0.948 (AOI 16), while AUC varied from
0.725 (AOI 4) and 0.93 (AOI 11), and F1 from 0.252 (AOI 13) and 0.636
(AOI 8).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figures/figure5}

}

\caption{Scores for overall accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 scores for the Random Forests model results after each iteration of the active learning loop for each AOI (gray lines), as well as the mean score per iteration across all AOIs (black lines).}\label{fig:alperformance}
\end{figure}

The comparison of active versus randomized training sample collection
(in AOIs 1, 8, and 15) showed that the former approach outperformed the
latter. After three iterations, the accuracy, AUC, and F1 scores
resulting from active learning were respectively 0.8, 0.6, and 2.3
percent higher than the scores from a randomly trained model (Figure
S8). However, there was more variability in earlier iterations, with
average score differences of -1.7 (accuracy), 0.6 (AUC), and 0.8 percent
(F1) after iteration 1, and -0.3 (accuracy), 0.4 (AUC), and 1.8 (F1)
percent after iteration 2. The negative results for accuracy was caused
by results at AOI 15, where active learning accuracy was 8.37 percent
lower than random training after iteration 1 (see Figure
\ref{fig:alperformance}). In comparison, iteration 1 active learning
accuracies were 2.88 and 0.45 percent higher than random training for
AOIs 1 and 8, respectively. Accuracy under active learning for AOI 15
exceeded randomized training after 4 iterations.

\hypertarget{the-impact-of-training-data-error}{%
\subsubsection{The impact of training data
error}\label{the-impact-of-training-data-error}}

The potential impact of label errors on map quality was assessed in four
AOIs (1, 2, 8, and 15). The results of these tests showed that the
average accuracy, AUC, and F1 scores for models trained with the
consensus labels were respectively 0.772, 0.8, and 0.555 (Figure
\ref{fig:trainingimpact}). Performance metrics from consensus-trained
models were just 0.5 - 1.2 percent higher than those models trained with
the most accurate individuals' labels (accuracy = 0.762; AUC = 0.796; F1
= 0.55), but were 11.6 - 27.4 higher than models trained with the least
accurate individual labels (accuracy = 0.606; AUC = 0.716; F1 = 0.44).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figures/figure6}

}

\caption{Scores for overall accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 score resulting from models trained with consensus labels, and labels made by the most and least accurate labellers to map each site. Comparisons were made for AOIs 1, 2, 8, and 15, denoted by grey symbols, while the mean scores across these AOIs are shown for each metric.}\label{fig:trainingimpact}
\end{figure}

A second measure of the impact of label error is found within the
correlations between the mean label risk per AOI and the model
performance metrics (Table S3). Accuracy and AUC had strong (Spearman's
Rank Correlation = -0.824 to moderate (r = -0.568) negative correlations
with label risk, while F1 had a weaker but moderate positive association
(r = 0.456). The positive sign of the latter relationship is
counter-intuitive, but is explained by risk's association with
precision, one of two inputs to F1, which was moderately positive
(0.629), whereas risk had a negligible correlation with recall (0.206),
F1's other component. The correlation between risk and the false
positive rate (0.688), another important performance metric, shows that
labelling uncertainty may increase model commission error.

\hypertarget{map-accuracy}{%
\subsection{Map accuracy}\label{map-accuracy}}

\hypertarget{categorical-accuracy}{%
\subsubsection{Categorical accuracy}\label{categorical-accuracy}}

We used a map reference sample of 1207 sites (487 cropland; 720
non-cropland) to evaluate the accuracy of the per-pixel classifications
(resulting from thresholding the Random Forests probability), as well as
the segmented field boundary maps. We first evaluated the uncertainty in
the map reference classes by assessing 1) the overall agreement between
map reference labels collected by two separate supervisors at 23 sites,
and 2) the confidence of the labels assigned by the supervisors (see SI
for details). The first measure showed that the two individual
supervisors' labels agreed at 87\% of common sites, while the second
showed that 15.7 of sites were labelled with the two classes that
indicated a level of uncertainty.

We found that the overall accuracy of the pixel-wise classifications was
88\% against this map reference sample (Table \ref{tab:mapaccuracy}).
Confining the map reference sample to four distinct zones (Figure S10A)
shows that overall accuracy ranged from 83.3\% in Zone 1 (AOIs 1-3) to
93.6\% in Zone 3 (AOIs 10, 11, 13, 15, and 16). The Producer's accuracy
of the cropland class was 61.7\% across Ghana, ranging from 45.6\% in
Zone 3 to 67.9\% in Zone 1, while the User's accuracy for was 67.3\%
overall, ranging from 59.8\% in Zone 4 to 71.2\% in Zone 1. Both
measures of accuracy were substantially higher for the non-cropland
class across all zones, typically exceeding 90\%. The lowest accuracies
for the non-cropland class was in Zone 1 (Producer's = 89.3\%; User's =
87.7\%).

The overall accuracies obtained from the segmented maps were generally
1-2 percentage points lower than those of the per-pixel maps, while
User's accuracies tended to be 8-10 percentage points less (Table
\ref{tab:mapaccuracy}). In contrast, Producer's accuracies were 15-20
points higher than in the per-pixel map. The segmentation step therefore
helped to reduce omission error while substantially increasing
commission error.

\begin{table}
\caption{Map accuracies and adjusted area estimates for the ~3 m pixel-wise classifications (based on Random Forests predictions; top 5 rows) and the segmented map (bottom 5 rows). Results are provided for 4 zones (Zone 1 = AOIs 1-3; Zone 2 = AOIs 4-9; Zone 3 = AOIs 10, 11, 13, 14, 16; Zone 4 = AOIs 12, 15) plus the entire country. The error matrix (with reference values in columns) provides the areal percentage for each cell, and the Producer's (P), User's (U), and overall (O) map accuracies and their margins of error (in parenthesis) are provided, as well as the sample-adjusted area estimates (in km$^{2}$) and margins of error. }
\includegraphics[width = 18cm]{figures/table2.png}
\label{tab:mapaccuracy}
\end{table}

\hypertarget{segmentation-accuracy}{%
\subsubsection{Segmentation accuracy}\label{segmentation-accuracy}}

The comparisons of digitized versus segmented field boundaries showed
that the mean field size across all validation sites averaged 4.97 ha
(Median = 3.75; StDev = 6.04), which was 1.41 times larger than the 2.06
ha (Median = 1.35; StDev = 3.26) mean area of labeller-digitized
polygons. This discrepancy was primarily caused by results in four AOIs
(2, 3, 7, and 15; Figure S11), where segments averaged between 7.76 and
10.76 ha, compared to 2.18 - 2.77 ha for the corresponding
hand-digitized polygons. The number of segmented fields per validation
site averaged 3.08 (median = 2.66; StDev = 2.9) compared to 4.4 (median
= 3.38; StDev = 4.52) for digitized polygons (Figure S12).

\hypertarget{ghanas-croplands}{%
\subsection{Ghana's croplands}\label{ghanas-croplands}}

Two separate maps of cropland were produced for each AOI, a per-pixel
map derived from the cropland probabilities, and the vectorized map of
field boundaries (Figure \ref{fig:mainmap}). The former provides the
more accurate picture of cropland distributions in Ghana, which are most
concentrated in the Southeastern corner (AOI 15), the central-western
region (AOI 7, the northeastern and northwestern corners of AOIs 10 and
11, and the south of AOI 8), and the northeastern quadrant stretching
from AOI 9 through AOIs 5 and 6 and up to AOIs 2 and 3. The northern
third of AOI 1 also has noticeable densities of cropland. Several
notable areas of low cropland density are also apparent, indicating the
presence of large protected areas, such as Mole National Park in the
southeastern corner of AOI 1 and Digya National Park in the northwestern
corner of AOI 12. In contrast, the relative absence of cropland in AOIs
13, 14, and 16 does not reflect the scarcity of agriculture in these
areas, but rather the predominance of tree crops, which we did not map.

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{figures/figure7}

}

\caption{The distribution of croplands in Ghana. The main map shows the percentage of croplands in each 0.005 degree grid cell, derived from the predicted cropland probabilities. The insets  on the margins illustrate predicted probabilities (top map in each couplet) at original image resolution (0.000025 degrees) and segmented field boundaries overlaid on the dry season PlanetScope composite, for four separate tiles. Each tile's position is shown on the main map, and is color-coded to the boundary lines around its corresponding inset.}\label{fig:mainmap}
\end{figure}

Both the per-pixel and vectorized maps, when combined with the map
reference sample, enable separate estimates of the total extent of
croplands in Ghana. The cropland extent estimated from the vectorized
map is 42,359 km\(^2\) (with a margin of error of 4,395 km\(^2\)), or
17.1 (15.4-18.9\%) of the mapped area. The estimate based on the per
pixel map is 43,233 km\(^2\) (margin of error = 4,904 km\(^2\)), or 17.6
(15.6-19.6\%) of area.

The vectorized map provides additional information on how the
characteristics of croplands can vary geographically, ranging from
narrow, strip-like fields in parts of AOI 15 (Figure \ref{fig:mainmap}'s
lower right inset) to more densely packed, less distinctly shaped fields
in AOI 5 (upper right inset in Figure \ref{fig:mainmap}). To explore how
field characteristics varied geographically, we mapped the average field
size and total number of fields within each 0.05 degree tile (Figure
S13). These patterns generally correspond to those seen in the cropland
density map (Figure \ref{fig:mainmap}), with larger sizes and field
counts generally occurring in areas of higher field density, although
the biases inherent in both measures (Figures S11-12) complicate the
interpretation of those variations. However, we can use the estimated
biases to develop adjusted estimates of field sizes and counts for each
AOI, and for Ghana overall (Table \ref{tab:sizentab}). These adjusted
estimates show that the typical field size in Ghana is 1.73 ha, ranging
from 0.96 in AOI 4 to 2.82 ha in AOI 4, with fields in the forest zone
AOIs (10, 11, 13, 14, 16) generally smaller than those in the northern
half of the country (Table \ref{tab:sizentab}). The total number of
fields was estimated to be 1,662,281 overall, or 205 fields per tile on
average, ranging from 108/tile in AOI 4 to 399/tile in AOI 6.

\begin{table}[!h]

\caption{\label{tab:sizentab}The average size and total number of crop fields for each AOI and for Ghana overall. The original and bias-adjusted values for each measure are provided, as well as the total number of 0.05$^\circ$ degree tiles in each AOI.}
\centering
\begin{tabular}[t]{>{\raggedleft\arraybackslash}p{0.45in}>{\raggedleft\arraybackslash}p{0.45in}>{\raggedleft\arraybackslash}p{0.45in}>{\raggedleft\arraybackslash}p{0.85in}>{\raggedleft\arraybackslash}p{0.85in}>{\raggedleft\arraybackslash}p{0.85in}>{\raggedleft\arraybackslash}p{0.85in}>{\raggedleft\arraybackslash}p{0.85in}}
\toprule
AOI & N tiles & Size & Size (adj) & N & N / tile & N (adj) & N (adj) / tile\\
\midrule
1 & 777 & 3.71 & 1.26 & 97,822 & 126 & 127,580 & 164\\
2 & 597 & 7.66 & 1.96 & 87,666 & 147 & 120,651 & 202\\
3 & 501 & 8.24 & 2.18 & 108,819 & 217 & 104,422 & 208\\
4 & 465 & 2.44 & 2.82 & 26,276 & 57 & 50,163 & 108\\
5 & 400 & 4.24 & 2.09 & 43,290 & 108 & 53,756 & 134\\
\addlinespace
6 & 429 & 5.10 & 2.15 & 81,363 & 190 & 145,347 & 339\\
7 & 471 & 5.64 & 1.49 & 93,282 & 198 & 123,005 & 261\\
8 & 400 & 4.89 & 1.98 & 55,500 & 139 & 78,868 & 197\\
9 & 479 & 4.10 & 1.82 & 72,081 & 150 & 89,840 & 188\\
10 & 630 & 2.24 & 1.04 & 119,019 & 189 & 170,907 & 271\\
\addlinespace
11 & 400 & 3.65 & 1.52 & 52,510 & 131 & 94,709 & 237\\
12 & 471 & 3.44 & 1.77 & 44,667 & 95 & 52,947 & 112\\
13 & 627 & 0.84 & 0.96 & 67,996 & 108 & 125,368 & 200\\
14 & 400 & 1.09 & 2.72 & 56,006 & 140 & 101,767 & 254\\
15 & 548 & 4.95 & 1.54 & 75,752 & 138 & 105,681 & 193\\
\addlinespace
16 & 521 & 0.95 & 1.41 & 49,097 & 94 & 117,268 & 225\\
Ghana & 8,116 & 3.92 & 1.73 & 1,131,146 & 139 & 1,662,281 & 205\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

These results demonstrate a capability for mapping the characteristics
of smallholder-dominated cropping systems at high spatial resolution,
annual time steps, and national scales. The resulting maps provide an
updated and more granular view of the distribution and extent of
croplands in Ghana, complementing existing national to regional land
cover maps derived from moderate resolution sensors (Hackman et al.
2017, Xiong et al. 2017, ESA n.d.). This prior work found that cropland
covered 19.4 (Xiong et al. 2017) to 32\% (Hackman et al. 2017) of Ghana
in 2015, whereas our 2018 maps have cropland cover of 16.1-23.2\% (Table
\ref{tab:mapaccuracy}), and our map reference sample-based estimates
finds 17.1-17.6\% cover. Our results thus suggest that Ghana's cropland
is less than previously estimated, but the difference is perhaps
attributable to our use of a cropland definition that excluded longer
fallows and abandoned fields, which in some regions can account for over
half of the total area that could be counted as cropland (Tong et al.
2020).

In addition to the more detailed update of cropland extent, our maps
also provide new information on the size and number of fields in Ghana
(Figures \ref{fig:mainmap}, S11-12). Previous work to estimate such
agricultural characteristics have often focused on farm, rather than
field, size using census data (Von Braun 2004, Samberg et al. 2016,
Jayne et al. 2016, Lowder et al. 2016). Efforts to map field boundaries
in smallholder-dominated agricultural systems have either used \emph{in
situ} data collection (Carletto et al. 2013, 2015) or remote sensing
studies over relatively small (e.g. Forkuor et al. 2014, Persello et al.
2019) or discontiguous (Estes et al. 2016a) areas. The most extensive
studies to date used crowdsourced volunteers to classify fields into
broad size classes, based on their interpretations of imagery sampled
from high resolution virtual globes (Fritz et al. 2015, Lesiv et al.
2019). Those efforts included country-specific results for Ghana (n =
263), which can be converted into an average field size estimate of 5.33
ha\footnote{Obtained by calculating the weighted mean from the count of the five size classes and the mean of the hectare range provided for the four smallest size classes, and the lower bound of the size range provided the largest size class. Data sourced from Table S3 in Lesiv et al. 2019.}.
This estimate exceeds our Ghana-wide average segment size (3.92 ha;
Table \ref{tab:sizentab}), but is closer to the mean (4.97 ha) within
AOIs 1-9, 12, and 15, which is where most of the crowdsourced sample
appears to have been collected. However, our bias-corrected estimates of
1.73 (Ghana-wide) and 1.87 (AOIs 1-9, 12, and 15) ha were much smaller.

\hypertarget{map-accuracy-and-key-sources-of-error}{%
\subsection{Map accuracy and key sources of
error}\label{map-accuracy-and-key-sources-of-error}}

Although the maps generated by our system provide valuable new
information, they nevertheless contain substantial errors. The overall
map accuracies (86.7-88\%, Table \ref{tab:mapaccuracy}) are near the
boundary of what might be considered \emph{achievable} map accuracy
(Elmes et al. 2020), given the inherent uncertainty in the map reference
sample, our best estimate of the ``truth,'' in which we have roughly
85\% confidence. However, accuracies for the cropland class were much
lower, falling between 62 (Producer's) to 67 (User's) percent
country-wide for the per-pixel map (Table \ref{tab:mapaccuracy}),
meaning the model produced substantial commission and omission errors
for this class. The segmented boundary maps had fewer omission errors
(Producer's accuracy = 79\%), but higher false positive errors (User's
accuracy = 58.2\%). These accuracies are near the middle to upper ranges
of those reported for the cropland class in other large-area mapping
studies (Hackman et al. 2017, Xiong et al. 2017, Lesiv et al. 2017).

The patterns of cropland-class accuracies varied by zone. These zones
largely align, albeit with some discrepancies, with the country's
agroecozones, thus the accuracy patterns may be partially attributed to
some regions being harder to map than others. Producer's accuracies for
both maps were highest in the two northern zones (1 and 2), which are
primarily savannas (Figure S10), and lowest in zones 3 and 4, which are
comprised of forest or coastal savannas. User's accuracies followed a
similar pattern, with the exception of Zone 3, which had the highest
User's accuracy, albeit from a very small sample. Aligning the reference
samples more precisely with agroecozone boundaries (Figure S10B)
provides further insight into error patterns (Table S4). Coastal
savannas in the southeast had the highest Producer's accuracy but lowest
User's accuracy for the per-pixel map, presumably because this region's
numerous areas of high density cropland, combined with low woody cover
in surrounding uncultivated areas, helped to promote commission error.
Maps in the two northern savanna agroecozones had the best balance
between omission and commission error, and had the highest overall
User's accuracy. The transitional (between forest and savanna)
agroecozone had a very low Producer's accuracy (21\%), which likely
reflects the fact that it was divided between several AOIs for mapping
(Figure S4), within which it typically covered a smaller share of area
relative to the other agroecozones. This likely caused insufficient
representation of this AEZ in training samples, particularly in AOIs 10
and 11 (Figure S4B).

Beyond the errors linked to regional differences, several other
important factors contributed to reducing the accuracies in the cropland
class. The first of these stems from the overall mapping extent and the
high resolution of the data. Given the goal of developing a
country-scale map at high resolution, the attendant data volume required
us to use a relatively small set of image features and less than the
recommended tree number and depth (Maxwell et al. 2018) in our Random
Forests implementation, in order to limit computational costs. Previous
work found that Random Forests achieves much better performance on
small-scale croplands when trained on a much larger number of features
(Debats et al. 2016). However, applying such a large feature set within
the extent of our AOIs was intractable, as the computing time would have
been several-fold larger than the \textasciitilde4-8 hours of runtime on
800 CPUs required for a single active learning iteration, followed by
\textasciitilde10-14 hours for prediction. This reduced the skill of the
model, particularly when it came to differentiating cropland from
adjacent bare patches or natural vegetation with sparse herbaceous
cover, which were common in many AOIs.

The inherent difficulty of the labelling task was another major limiting
factor. Our system was designed to minimize the error inherent in
labelling, but determining croplands from non-croplands in these
agricultural systems can be a difficult task. Labellers have to evaluate
multiple image sources and to rely heavily on judgement, which
inevitably leads to errors. Interpretation is particularly hard where
the background savanna vegetation and croplands have similar reflectance
during the dry season, which is a particular problem in AOIs 2 and 3.
Smaller field sizes also complicate labelling, as these become
increasingly indistinct in the \textasciitilde4 m PlanetScope
composites. The difficulty of labelling is reflected in the magnitude of
the Bayesian Risk metrics (Figure S6), and by the average score achieved
by each labeller against our training reference dataset (71\%; Figure
S5B). Although prior work (Rodriguez-Galiano et al. 2012, Mellor et al.
2015) found that Random Forests are robust to label errors, we found
that they have substantial impact (Figure \ref{fig:trainingimpact}),
which suggest that simply improving label quality may be one of the
single most important investments towards improved model accuracy.

Image quality was another issue, although primarily in the forested
AOIs, where frequent cloud cover reduced the number of available images
in all seasons, resulting in composites with more brightness artifacts
and blur (Figure \ref{fig:imqual}). This impacted labellers' abilities
to discern fields, and doubtless affected model predictions. There is
little to be done to mitigate these errors, short of confining imagery
to the less cloudy dry season, which may further undermine model
performance, given the importance of multi-temporal imagery for cropland
classification (Debats et al. 2016, Defourny et al. 2019). Composite
quality could be improved by using imagery from the same seasons over
multiple years, but this would undermine the goal of developing annual
maps, while the dynamism of the croplands would blur field boundaries
within the imagery.

The final major source of error arose from the segmentation process. The
vectorized maps had high commission errors caused by uncertainties in
the Random Forests predictions. Model uncertainty led to many pixels
over non-crop areas with probabilities straddling the 0.5 classification
threshold. Segments that intersected such areas were retained as fields
when the average probability of intersecting pixels exceeded 0.5. A more
accurate classifier would reduce such errors, or the application of a
locally varying classification threshold (e.g. Waldner and Diakogiannis
2020). Over-merging was another source of error in the segmentation
algorithm, which in some areas led to overestimated field sizes and
unrealistic shapes, particularly in high density croplands (e.g.~in AOIs
2 and 8; Figure \ref{fig:mainmap}) where the boundaries between adjacent
fields are often indistinct in the PlanetScope imagery. Minimizing or
preventing merging would help in such cases, although could would result
in the opposite problem, over-segmentation, and thereby underestimate
field size.

\hypertarget{error-mitigation-features}{%
\subsection{Error mitigation features}\label{error-mitigation-features}}

Despite the error sources mentioned above, several features of the
system proved effective in mitigating error, leading to a higher overall
accuracy than would have otherwise been possible. Label accuracy
assessment and consensus labelling appeared to be the most effective
error mitigation tools. Label accuracy measures allowed us to quantify
the substantial impact of label error on model performance (Figure
\ref{fig:trainingimpact}), while consensus labels substantially reduced
individual labelling errors, resulting in maps that were more accurate
than they would have been had we used individually generated labels.
Labeller-specific accuracy measures also helped to improve the quality
of the consensus labels, by placing higher weight on labels more likely
to be accurate during the merging process, rather than giving equal
weight to potentially less accurate labels. The ability to select the
most accurate individual labels for a site also allowed us to develop
independent estimates of field size to which measures of confidence can
be attached (Figure S5B), which we were in turn able to correct
estimates of field sizes and numbers (Table \ref{tab:sizentab}).

The active learning approach helped to improve overall model performance
relative to randomized training site selection, in line with findings
from two recent efforts (Debats et al. 2017, Hamrouni et al. 2021).
Although the performance gains relative to randomized model training
that we observed were smaller (e.g. Debats et al. (2017) 29\% higher
model performance after one iteration, and 8\% higher on the final
iterations), those comparisons were made from lower initial bases, with
initial training samples that were less than 1/10 the size, in terms of
pixels, of our initial training sample. Our large initial randomly
selected sample (500 grid cells) meant that our models were
substantially trained before they were exposed to actively selected
labels, thereby diluting their impact on performance. Nevertheless, the
higher average performance of the active approach across three
performance metrics demonstrated its effectiveness. Most notable were
the larger improvements seen in the F1 score (Figure S8), a balanced
performance metric. Gains in accuracy and AUC were smaller. For
accuracy, the reduced advantage was primarily due to active learning
being outperformed by randomized training after the first iteration in
AOI 15, which proved one of the hardest AOIs to both map and label.
Active learning likely resulted in the selection of sites that were
harder to label than randomly selected ones, leading to more label
error, and thus initially lower model accuracy. However, this deficit
was overcome by the 5th iteration. The plateau in AUC gains at 0.5\%
better than randomized training reflects the findings that active
learning reduced both the false and true positive rates, the two inputs
to AUC. Although the decline of the false positive rate (30.7\% between
Iterations 0 and 3) was nearly three times larger than that of the true
positive rate (10.9\%), AUC should be quite sensitive to the reduction
in the latter, as it assesses how the tradeoff between the two rates
varies across a full range of possible classification thresholds
(Pontius and Si 2014).

The detail, temporal precision, and large extent of these maps was
enabled by the system's ability to process PlanetScope data, which is
currently the only source of sub-5 meter imagery with daily coverage
(McCabe et al. 2017). Daily revisits are important for creating seasonal
composites within a single year over cloudy areas. The compositing
technique we developed allowed us to develop a complete image catalog
for the country representing the two seasons for 2018 agricultural year.
Although Sentinel-2 is free, has better radiometric quality, and has
sufficiently high resolution (10 m) to accurately classify small-scale
agricultural systems (e.g. Defourny et al. 2019, Kerner et al. 2020),
its 5-day interval may be too infrequent to generate sufficiently
cloud-free composites during the growing season over southern Ghana.
Sentinel-1 is not affected by the same problem, but labelling fields in
more coarsely resolved radar images is challenging.

\hypertarget{lingering-questions}{%
\subsection{Lingering questions}\label{lingering-questions}}

Several potential issues not addressed in our assessment merit further
exploration. One of these is the degree of correspondence between image-
and ground-collected labels. However, such comparisons may reveal
unresolvable differences between the two perspectives. The highly
dynamic nature of many agricultural systems means that relatively narrow
differences between the dates of ground- and image-based digitizing
campaigns can lead to substantial disagreement between the resulting
field boundaries, simply because the fields themselves may have shifted
during the interval (Elmes et al. 2020). These discrepancies could be
further exacerbated by differences in the definition of what constitutes
a field, which might vary on the ground depending on who is being asked,
or who is doing the collecting. These factors suggest that ground versus
image label differences would not necessarily indicate how far
image-based labellers were from the ``truth.'' Nevertheless, a
comparison against ground data would help to assess how accurately
image-collected labels capture the typical size of fields, and thus
merits further investigation for this reason.

The temporal discrepancies mentioned above (and discussed in Elmes et
al. 2020) are another reason why we chose not to label on basemap
imagery (in addition to restrictive usage terms), which is typically
several years old (Lesiv et al. 2018). However, we did not assess
whether the higher label accuracy one might achieve by digitizing on a
\textless1-2 m resolution basemap would offset model errors caused by
temporal mismatches.

Another potential issue is the degree to which our assessment of the
impact of label error on model performance (Figure
\ref{fig:trainingimpact}) was influenced by the validation labels we
used, which were generated using the consensus method. This could have
confounded the assessment, particularly when comparing models trained
with the most accurate individual label and those trained with consensus
labels. However, the visual assessment of their resulting probability
maps confirm the differences in scores: consensus and most accurate
individual labels produce nearly identical maps with relatively high
certainty, while low quality labels led to a markedly less certain map
(Figure S9).

\hypertarget{next-steps}{%
\subsection{Next steps}\label{next-steps}}

The maps presented here represent a version 1 product that is freely
available to use, along with its underlying code (see SI for details).
These data were developed according to the recommended best practices
for training and assessing error in machine learning models (Elmes et
al. 2020). In their current form, the maps may be useful for a variety
of research applications. For example, analyzing the distributions of
values in the probability maps may provide additional insight into the
relative extents of active versus fallow croplands (Tong et al. 2020).
However, use of these data, particularly for decision-making processes
(e.g.~cropped area estimates), should be careful to account for the
reported errors (Olofsson et al. 2014, Stehman and Foody 2019).

To facilitate the next step, generating more accurate version 2 maps,
several improvements will be made. The first is to replace Random
Forests with a more advanced convolutional neural network (CNN), which
can generate and learn from a large number of features representing a
variety of spatial scales (Ma et al. 2019). Recent work suggests that a
common architecture such as U-Net, when trained to distinguish field
edges from interiors and combined with a post-hoc segmentation routine,
is effective in delineating field boundaries (Waldner and Diakogiannis
2020). Our system can readily incorporate such a model. The labelling
platform already provides the methods needed to develop and assess the
quality of labels that include field edges and interior classes, while
active learning has proven to be effective for optimizing training
datasets for deep learning models (Liu et al. 2017, Cao et al. 2020).
Our current framework can be adjusted so that it starts by training a
CNN from scratch with a large initial random sample, and then uses a
transfer learning approach (Pan and Yang 2010) to update the model with
the most informative samples from different AOIs or agroecozones.

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

This work demonstrates a proof of concept for developing high
resolution, annual maps of smallholder-dominated croplands at national
to regional scales, using a framework that can be readily updated to
improve map accuracy as technologies improve. Maps that include
information on field boundaries can help improve remote estimation of
crop planted area and yield, and provide deeper insights into important
socioeconomic aspects of agricultural systems, such as the relationships
between agricultural productivity and livelihoods. Such maps will be
important for developing an understanding of the rapid agricultural
change that is currently unfolding throughout much of the continent.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

The primary support for this work was provided by Omidyar Network's
Property Rights Initiative, now PLACE. Additional support was provided
by NASA (80NSSC18K0158), the National Science Foundation (SES-1801251;
SES-1832393), and Princeton University. Computing support was provided
by the AWS Cloud Credits for Research program and the Amazon
Sustainability Data Initiative. Azavea provided significant
contributions in engineering the machine learning pipeline. We thank
Meridia for providing information about local cropping systems and the
characteristics of fields, and Radiant Earth Foundation for advice and
guidance regarding machine learning best practices. We thank Manushi
Trivedi, Sitian Xiong, and Tammy Woodard for their contributions to the
underlying datasets and methods, and Michelle Gathigi, Omar Shehe, and
Primoz Kovacic for support and management of the labelling efforts.

\hypertarget{references}{%
\section{References}\label{references}}

\singlespace

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-azaveaRasterFoundry2020}{}%
Azavea. 2020. Raster {Foundry}.
https://github.com/raster-foundry/raster-foundry.

\leavevmode\hypertarget{ref-BeyCollectEarthLand2016}{}%
Bey, A., A. Sánchez-Paus Díaz, D. Maniatis, G. Marchi, D. Mollicone, S.
Ricci, J.-F. Bastin, R. Moore, S. Federici, M. Rezende, C. Patriarca, R.
Turia, G. Gamoga, H. Abe, E. Kaidong, and G. Miceli. 2016. Collect
{Earth}: {Land Use} and {Land Cover Assessment} through {Augmented
Visual Interpretation}. Remote Sensing 8:807.

\leavevmode\hypertarget{ref-BreimanRandomForests2001}{}%
Breiman, L. 2001. Random {Forests}. Machine Learning 45:5--32.

\leavevmode\hypertarget{ref-bullockThreeDecadesLand2021}{}%
Bullock, E. L., S. P. Healey, Z. Yang, P. Oduor, N. Gorelick, S. Omondi,
E. Ouko, and W. B. Cohen. 2021. Three {Decades} of {Land Cover Change}
in {East Africa}. Land 10:150.

\leavevmode\hypertarget{ref-caoHyperspectralImageClassification2020}{}%
Cao, X., J. Yao, Z. Xu, and D. Meng. 2020. Hyperspectral {Image
Classification With Convolutional Neural Network} and {Active Learning}.
IEEE Transactions on Geoscience and Remote Sensing 58:4604--4616.

\leavevmode\hypertarget{ref-CarlettoGuesstimatesGPStimatesLand2015}{}%
Carletto, C., S. Gourlay, and P. Winters. 2015. From {Guesstimates} to
{GPStimates}: {Land Area Measurement} and {Implications} for
{Agricultural Analysis}. Journal of African Economies 24:593--628.

\leavevmode\hypertarget{ref-CarlettoFactartifactimpact2013}{}%
Carletto, C., S. Savastano, and A. Zezza. 2013. Fact or artifact: {The}
impact of measurement errors on the farm size{}productivity
relationship. Journal of Development Economics 103:254--261.

\leavevmode\hypertarget{ref-ChenGloballandcover2015}{}%
Chen, J., J. Chen, A. Liao, X. Cao, L. Chen, X. Chen, C. He, G. Han, S.
Peng, M. Lu, W. Zhang, X. Tong, and J. Mills. 2015. Global land cover
mapping at 30 m resolution: {A POK}-based operational approach. ISPRS
Journal of Photogrammetry and Remote Sensing 103:7--27.

\leavevmode\hypertarget{ref-cohnImprovingGeneralizationActive1994}{}%
Cohn, D., L. Atlas, and R. Ladner. 1994. Improving generalization with
active learning. Machine Learning 15:201--221.

\leavevmode\hypertarget{ref-Darkmodifiablearealunit2007}{}%
Dark, S. J., and D. Bram. 2007. The modifiable areal unit problem
({MAUP}) in physical geography. Progress in Physical Geography
31:471--479.

\leavevmode\hypertarget{ref-davisTropicalForestLoss2020}{}%
Davis, K. F., H. I. Koo, J. Dell'Angelo, P. D'Odorico, L. Estes, L. J.
Kehoe, M. Kharratzadeh, T. Kuemmerle, D. Machava, A. de J. R. Pais, N.
Ribeiro, M. C. Rulli, and M. Tatlhego. 2020. Tropical forest loss
enhanced by large-scale land acquisitions. Nature Geoscience:1--7.

\leavevmode\hypertarget{ref-DebatsIntegratingactivelearning2017}{}%
Debats, S. R., L. D. Estes, D. R. Thompson, and K. K. Caylor. 2017.
Integrating active learning and crowdsourcing into large-scale
supervised landcover mapping algorithms. {PeerJ Preprints}.

\leavevmode\hypertarget{ref-Debatsgeneralizedcomputervision2016}{}%
Debats, S. R., D. Luo, L. D. Estes, T. J. Fuchs, and K. K. Caylor. 2016.
A generalized computer vision approach to mapping crop fields in
heterogeneous agricultural landscapes. Remote Sensing of Environment
179:210--221.

\leavevmode\hypertarget{ref-Defournyrealtimeagriculturemonitoring2019}{}%
Defourny, P., S. Bontemps, N. Bellemans, C. Cara, G. Dedieu, E.
Guzzonato, O. Hagolle, J. Inglada, L. Nicola, T. Rabaute, M. Savinaud,
C. Udroiu, S. Valero, A. Bégué, J.-F. Dejoux, A. El Harti, J. Ezzahar,
N. Kussul, K. Labbassi, V. Lebourgeois, Z. Miao, T. Newby, A. Nyamugama,
N. Salh, A. Shelestov, V. Simonneaux, P. S. Traore, S. S. Traore, and B.
Koetz. 2019. Near real-time agriculture monitoring at national scale at
parcel resolution: {Performance} assessment of the {Sen2}-{Agri}
automated system in various cropping systems around the world. Remote
Sensing of Environment 221:551--568.

\leavevmode\hypertarget{ref-desiereLandProductivityPlot2018}{}%
Desiere, S., and D. Jolliffe. 2018. Land productivity and plot size:
{Is} measurement error driving the inverse relationship? Journal of
Development Economics 130:84--98.

\leavevmode\hypertarget{ref-DruschSentinel2ESAOptical2012}{}%
Drusch, M., U. Del Bello, S. Carlier, O. Colin, V. Fernandez, F. Gascon,
B. Hoersch, C. Isola, P. Laberinti, P. Martimort, A. Meygret, F. Spoto,
O. Sy, F. Marchese, and P. Bargellini. 2012. Sentinel-2: {ESA}'s
{Optical High}-{Resolution Mission} for {GMES Operational Services}.
Remote Sensing of Environment 120:25--36.

\leavevmode\hypertarget{ref-DwyerAnalysisReadyData2018}{}%
Dwyer, J. L., D. P. Roy, B. Sauer, C. B. Jenkerson, H. K. Zhang, and L.
Lymburner. 2018. Analysis {Ready Data}: {Enabling Analysis} of the
{Landsat Archive}. Remote Sensing 10:1363.

\leavevmode\hypertarget{ref-ElmesAccountingtrainingdata2020}{}%
Elmes, A., H. Alemohammad, R. Avery, K. Caylor, J. R. Eastman, L.
Fishgold, M. A. Friedl, M. Jain, D. Kohli, J. C. Laso Bayas, D. Lunga,
J. L. McCarty, R. G. Pontius, A. B. Reinmann, J. Rogan, L. Song, H.
Stoynova, S. Ye, Z.-F. Yi, and L. Estes. 2020. Accounting for training
data error in machine learning applied to {Earth Observations}. Remote
Sensing 12:1034.

\leavevmode\hypertarget{ref-ESAESACCILAND}{}%
ESA. (n.d.). {ESA CCI LAND COVER} {} {S2} prototype {Land Cover} 20m map
of {Africa} 2016. http://2016africalandcover20m.esrin.esa.int/.

\leavevmode\hypertarget{ref-Esteslargeareaspatiallycontinuous2018}{}%
Estes, L., P. Chen, S. Debats, T. Evans, S. Ferreira, T. Kuemmerle, G.
Ragazzo, J. Sheffield, A. Wolf, E. Wood, and K. Caylor. 2018. A
large-area, spatially continuous assessment of land cover map error and
its impact on downstream analyses. Global Change Biology 24:322--337.

\leavevmode\hypertarget{ref-Estesplatformcrowdsourcingcreation2016}{}%
Estes, L. D., D. McRitchie, J. Choi, S. Debats, T. Evans, W. Guthe, D.
Luo, G. Ragazzo, R. Zempleni, and K. K. Caylor. 2016a. A platform for
crowdsourcing the creation of representative, accurate landcover maps.
Environmental Modelling \& Software 80:41--53.

\leavevmode\hypertarget{ref-EstesReconcilingagriculturecarbon2016}{}%
Estes, L. D., T. Searchinger, M. Spiegel, D. Tian, S. Sichinga, M.
Mwale, L. Kehoe, T. Kuemmerle, A. Berven, N. Chaney, J. Sheffield, E. F.
Wood, and K. K. Caylor. 2016b. Reconciling agriculture, carbon and
biodiversity in a savannah transformation frontier. Phil. Trans. R. Soc.
B 371:20150316.

\leavevmode\hypertarget{ref-federRelationFarmSize1985}{}%
Feder, G. 1985. The relation between farm size and farm productivity:
{The} role of family labor, supervision and credit constraints. Journal
of Development Economics 18:297--313.

\leavevmode\hypertarget{ref-forkuorIntegrationOpticalSynthetic2014}{}%
Forkuor, G., C. Conrad, M. Thiel, T. Ullmann, and E. Zoungrana. 2014.
Integration of {Optical} and {Synthetic Aperture Radar Imagery} for
{Improving Crop Mapping} in {Northwestern Benin}, {West Africa}. Remote
Sensing 6:6472--6499.

\leavevmode\hypertarget{ref-FourieBetterCropEstimates2009}{}%
Fourie, A. 2009. Better {Crop Estimates} in {South Africa}. ArcUser
Online.

\leavevmode\hypertarget{ref-FritzGeoWikionlineplatform2012}{}%
Fritz, S., I. McCallum, C. Schill, C. Perger, L. See, D. Schepaschenko,
M. van der Velde, F. Kraxner, and M. Obersteiner. 2012. Geo-{Wiki}: {An}
online platform for improving global land cover. Environmental Modelling
\& Software 31:110--123.

\leavevmode\hypertarget{ref-FritzHighlightingcontinueduncertainty2011}{}%
Fritz, S., L. See, I. McCallum, C. Schill, M. Obersteiner, M. van der
Velde, H. Boettcher, P. Havlík, and F. Achard. 2011. Highlighting
continued uncertainty in global land cover maps for the user community.
Environmental Research Letters 6:044005.

\leavevmode\hypertarget{ref-FritzMappingglobalcropland2015}{}%
Fritz, S., L. See, I. McCallum, L. You, A. Bun, E. Moltchanova, M.
Duerauer, F. Albrecht, C. Schill, C. Perger, P. Havlik, A. Mosnier, P.
Thornton, U. Wood-Sichra, M. Herrero, I. Becker-Reshef, C. Justice, M.
Hansen, P. Gong, S. Abdel Aziz, A. Cipriani, R. Cumani, G. Cecchi, G.
Conchedda, S. Ferreira, A. Gomez, M. Haffani, F. Kayitakire, J.
Malanding, R. Mueller, T. Newby, A. Nonguierma, A. Olusegun, S. Ortner,
D. R. Rajak, J. Rocha, D. Schepaschenko, M. Schepaschenko, A. Terekhov,
A. Tiangwa, C. Vancutsem, E. Vintrou, W. Wenbin, M. van der Velde, A.
Dunwoody, F. Kraxner, and M. Obersteiner. 2015. Mapping global cropland
and field size. Global Change Biology 21:1980--1992.

\leavevmode\hypertarget{ref-FritzComparisonglobalregional2010}{}%
Fritz, S., L. See, and F. Rembold. 2010. Comparison of global and
regional land cover maps with statistical information for the
agricultural domain in {Africa}. International Journal of Remote Sensing
31:2237--2256.

\leavevmode\hypertarget{ref-Fritzneedimprovedmaps2013}{}%
Fritz, S., L. See, L. You, C. Justice, I. Becker-Reshef, L. Bydekerke,
R. Cumani, P. Defourny, K. Erb, J. Foley, S. Gilliams, P. Gong, M.
Hansen, T. Hertel, M. Herold, M. Herrero, F. Kayitakire, J. Latham, O.
Leo, I. McCallum, M. Obersteiner, N. Ramankutty, J. Rocha, H. Tang, P.
Thornton, C. Vancutsem, M. van der Velde, S. Wood, and C. Woodcock.
2013. The need for improved maps of global cropland. Eos, Transactions
American Geophysical Union 94:31--32.

\leavevmode\hypertarget{ref-GorelickGoogleEarthEngine2017}{}%
Gorelick, N., M. Hancher, M. Dixon, S. Ilyushchenko, D. Thau, and R.
Moore. 2017. Google {Earth Engine}: {Planetary}-scale geospatial
analysis for everyone. Remote Sensing of Environment 202:18--27.

\leavevmode\hypertarget{ref-hackmanNewLandcoverMaps2017}{}%
Hackman, K. O., P. Gong, and J. Wang. 2017. New land-cover maps of
{Ghana} for 2015 using {Landsat} 8 and three popular classifiers for
biodiversity assessment. International Journal of Remote Sensing
38:4008--4021.

\leavevmode\hypertarget{ref-hamrouniLocalGlobalTransfer2021}{}%
Hamrouni, Y., E. Paillassa, V. Chéret, C. Monteil, and D. Sheeren. 2021.
From local to global: {A} transfer learning-based approach for mapping
poplar plantations at national scale using {Sentinel}-2. ISPRS Journal
of Photogrammetry and Remote Sensing 171:76--100.

\leavevmode\hypertarget{ref-HouborgDailyRetrievalNDVI2018}{}%
Houborg, R., and M. McCabe. 2018. Daily {Retrieval} of {NDVI} and {LAI}
at 3 m {Resolution} via the {Fusion} of {CubeSat}, {Landsat}, and {MODIS
Data}. Remote Sensing 10:890.

\leavevmode\hypertarget{ref-jayneAfricaChangingFarm2016}{}%
Jayne, T. s., J. Chamberlin, L. Traub, N. Sitko, M. Muyanga, F. K.
Yeboah, W. Anseeuw, A. Chapoto, A. Wineman, C. Nkonde, and R. Kachule.
2016. Africa's changing farm size distribution patterns: The rise of
medium-scale farms. Agricultural Economics 47:197--214.

\leavevmode\hypertarget{ref-KansangaTraditionalagriculturetransition2019}{}%
Kansanga, M., P. Andersen, D. Kpienbaareh, S. Mason-Renton, K. Atuoye,
Y. Sano, R. Antabe, and I. Luginaah. 2019. Traditional agriculture in
transition: Examining the impacts of agricultural modernization on
smallholder farming in {Ghana} under the new {Green Revolution}.
International Journal of Sustainable Development \& World Ecology
26:11--24.

\leavevmode\hypertarget{ref-kehoeNatureRiskFuture2017}{}%
Kehoe, L., A. Romero-Muñoz, L. Estes, H. Kreft, E. Polaina, and T.
Kuemmerle. 2017. Nature at risk under future agricultural expansion and
intensification. Nature Ecology and Evolution 1:1129--1135.

\leavevmode\hypertarget{ref-kernerRapidResponseCrop2020}{}%
Kerner, H., G. Tseng, I. Becker-Reshef, C. Nakalembe, B. Barker, B.
Munshell, M. Paliyam, and M. Hosseini. 2020. Rapid {Response Crop Maps}
in {Data Sparse Regions}. arXiv:2006.16866 {[}cs, eess{]}.

\leavevmode\hypertarget{ref-LesivEvaluationESACCI2017}{}%
Lesiv, M., S. Fritz, I. McCallum, N. Tsendbazar, M. Herold, J.-F. Pekel,
M. Buchhorn, B. Smets, and R. Van De Kerchove. 2017, November.
Evaluation of {ESA CCI} prototype land cover map at 20m. Monograph,
http://pure.iiasa.ac.at/id/eprint/14979/.

\leavevmode\hypertarget{ref-lesivEstimatingGlobalDistribution2019}{}%
Lesiv, M., J. C. Laso Bayas, L. See, M. Duerauer, D. Dahlia, N. Durando,
R. Hazarika, P. Kumar Sahariah, M. Vakolyuk, V. Blyshchyk, A. Bilous, A.
Perez-Hoyos, S. Gengler, R. Prestele, S. Bilous, I. ul H. Akhtar, K.
Singha, S. B. Choudhury, T. Chetri, Ž. Malek, K. Bungnamei, A. Saikia,
D. Sahariah, W. Narzary, O. Danylo, T. Sturn, M. Karner, I. McCallum, D.
Schepaschenko, E. Moltchanova, D. Fraisl, I. Moorthy, and S. Fritz.
2019. Estimating the global distribution of field size using
crowdsourcing. Global Change Biology 25:174--186.

\leavevmode\hypertarget{ref-LesivCharacterizingspatialtemporal2018}{}%
Lesiv, M., L. See, J. Laso Bayas, T. Sturn, D. Schepaschenko, M. Karner,
I. Moorthy, I. McCallum, and S. Fritz. 2018. Characterizing the spatial
and temporal availability of very high resolution satellite imagery in
{Google Earth} and {Microsoft Bing} maps as a source of reference data.
Land 7:118.

\leavevmode\hypertarget{ref-levinFarmSizeLandscape2006}{}%
Levin, G. 2006. Farm size and landscape composition in relation to
landscape changes in {Denmark}. Geografisk Tidsskrift-Danish Journal of
Geography 106:45--59.

\leavevmode\hypertarget{ref-LickerMindgaphow2010}{}%
Licker, R., M. Johnston, J. A. Foley, C. Barford, C. J. Kucharik, C.
Monfreda, and N. Ramankutty. 2010. Mind the gap: How do climate and
agricultural management explain the {`yield gap'} of croplands around
the world? Global Ecology and Biogeography 19:769--782.

\leavevmode\hypertarget{ref-liuActiveDeepLearning2017}{}%
Liu, P., H. Zhang, and K. B. Eom. 2017. Active {Deep Learning} for
{Classification} of {Hyperspectral Images}. IEEE Journal of Selected
Topics in Applied Earth Observations and Remote Sensing 10:712--724.

\leavevmode\hypertarget{ref-lobellCropYieldGaps2009}{}%
Lobell, D. B., K. G. Cassman, and C. B. Field. 2009. Crop {Yield Gaps}:
{Their Importance}, {Magnitudes}, and {Causes}. Annual Review of
Environment and Resources 34:179--204.

\leavevmode\hypertarget{ref-LowderNumberSizeDistribution2016}{}%
Lowder, S. K., J. Skoet, and T. Raney. 2016. The {Number}, {Size}, and
{Distribution} of {Farms}, {Smallholder Farms}, and {Family Farms
Worldwide}. World Development 87:16--29.

\leavevmode\hypertarget{ref-maDeepLearningRemote2019}{}%
Ma, L., Y. Liu, X. Zhang, Y. Ye, G. Yin, and B. A. Johnson. 2019. Deep
learning in remote sensing applications: {A} meta-analysis and review.
ISPRS Journal of Photogrammetry and Remote Sensing 152:166--177.

\leavevmode\hypertarget{ref-MaxwellImplementationmachinelearningclassification2018}{}%
Maxwell, A. E., T. A. Warner, and F. Fang. 2018. Implementation of
machine-learning classification in remote sensing: An applied review.
International Journal of Remote Sensing 39:2784--2817.

\leavevmode\hypertarget{ref-McCabefutureEarthobservation2017}{}%
McCabe, M. F., M. Rodell, D. E. Alsdorf, D. G. Miralles, R. Uijlenhoet,
W. Wagner, A. Lucieer, R. Houborg, N. E. C. Verhoest, T. E. Franz, J.
Shi, H. Gao, and E. F. Wood. 2017. The future of {Earth} observation in
hydrology. Hydrology and Earth System Sciences 21:3879--3914.

\leavevmode\hypertarget{ref-mellorExploringIssuesTraining2015}{}%
Mellor, A., S. Boukir, A. Haywood, and S. Jones. 2015. Exploring issues
of training data imbalance and mislabelling on random forest performance
for large area land cover classification using the ensemble margin.
ISPRS Journal of Photogrammetry and Remote Sensing 105:155--168.

\leavevmode\hypertarget{ref-morrisAwakeningAfricaSleeping2009}{}%
Morris, M., and D. Byerlee. 2009. Awakening {Africa}'s {Sleeping Giant}.
{World Bank and FAO}, {Washington, DC}.

\leavevmode\hypertarget{ref-muellerClosingYieldGaps2012}{}%
Mueller, N. D., J. S. Gerber, M. Johnston, D. K. Ray, N. Ramankutty, and
J. A. Foley. 2012. Closing yield gaps through nutrient and water
management. Nature 490:254--257.

\leavevmode\hypertarget{ref-neubertCompactWatershedPreemptive2014}{}%
Neubert, P., and P. Protzel. 2014. Compact {Watershed} and {Preemptive
SLIC}: {On Improving Trade}-offs of {Superpixel Segmentation
Algorithms}. Pages 996--1001 2014 22nd {International Conference} on
{Pattern Recognition}. {IEEE}, {Stockholm, Sweden}.

\leavevmode\hypertarget{ref-OlofssonGoodpracticesestimating2014}{}%
Olofsson, P., G. M. Foody, M. Herold, S. V. Stehman, C. E. Woodcock, and
M. A. Wulder. 2014. Good practices for estimating area and assessing
accuracy of land change. Remote Sensing of Environment 148:42--57.

\leavevmode\hypertarget{ref-OlofssonMakingbetteruse2013}{}%
Olofsson, P., G. M. Foody, S. V. Stehman, and C. E. Woodcock. 2013.
Making better use of accuracy data in land change studies: {Estimating}
accuracy and area and quantifying uncertainty using stratified
estimation. Remote Sensing of Environment 129:122--131.

\leavevmode\hypertarget{ref-Openshawmillioncorrelationcoefficients1979}{}%
Openshaw, S., and P. J. Taylor. 1979. A million or so correlation
coefficients: Three experiments on the modifiable areal unit problem.
Statistical applications in the spatial sciences 21:127--144.

\leavevmode\hypertarget{ref-panSurveyTransferLearning2010}{}%
Pan, S. J., and Q. Yang. 2010. A {Survey} on {Transfer Learning}. IEEE
Transactions on Knowledge and Data Engineering 22:1345--1359.

\leavevmode\hypertarget{ref-perselloDelineationAgriculturalFields2019}{}%
Persello, C., V. A. Tolpekin, J. R. Bergado, and R. A. de By. 2019.
Delineation of agricultural fields in smallholder farms from satellite
images using fully convolutional networks and combinatorial grouping.
Remote Sensing of Environment 231:111253.

\leavevmode\hypertarget{ref-planetteamPlanetApplicationProgram2018}{}%
PlanetTeam. 2018. Planet application program interface: In space for
life on {Earth}. {https://api.planet.com}, {San Francisco, CA}.

\leavevmode\hypertarget{ref-Pontiustotaloperatingcharacteristic2014}{}%
Pontius, R. G., and K. Si. 2014. The total operating characteristic to
measure diagnostic ability for multiple thresholds. International
Journal of Geographical Information Science 28:570--583.

\leavevmode\hypertarget{ref-qiuCirrusCloudsThat2020}{}%
Qiu, S., Z. Zhu, and C. E. Woodcock. 2020. Cirrus clouds that adversely
affect {Landsat} 8 images: {What} are they and how to detect them?
Remote Sensing of Environment 246:111884.

\leavevmode\hypertarget{ref-rodriguez-galianoAssessmentEffectivenessRandom2012}{}%
Rodriguez-Galiano, V. F., B. Ghimire, J. Rogan, M. Chica-Olmo, and J. P.
Rigol-Sanchez. 2012. An assessment of the effectiveness of a random
forest classifier for land-cover classification. ISPRS Journal of
Photogrammetry and Remote Sensing 67:93--104.

\leavevmode\hypertarget{ref-RulliFoodappropriationlarge2014}{}%
Rulli, M. C., and P. D'Odorico. 2014. Food appropriation through large
scale land acquisitions. Environmental Research Letters 9:064030.

\leavevmode\hypertarget{ref-SambergSubnationaldistributionaverage2016}{}%
Samberg, L. H., J. S. Gerber, N. Ramankutty, M. Herrero, and P. C. West.
2016. Subnational distribution of average farm size and smallholder
contributions to global food production. Environmental Research Letters
11:124010.

\leavevmode\hypertarget{ref-SearchingerHighcarbonbiodiversity2015}{}%
Searchinger, T. D., L. Estes, P. K. Thornton, T. Beringer, A.
Notenbaert, D. Rubenstein, R. Heimlich, R. Licker, and M. Herrero. 2015.
High carbon and biodiversity costs from converting {Africa}'s wet
savannahs to cropland. Nature Climate Change 5:481--486.

\leavevmode\hypertarget{ref-searchingerCreatingSustainableFood2019}{}%
Searchinger, T., R. Waite, C. Hanson, J. Ranganathan, P. Dumas, E.
Matthews, and C. Klirs. 2019. Creating a sustainable food future: {A}
menu of solutions to feed nearly 10 billion people by 2050. {Final}
report. {WRI}.

\leavevmode\hypertarget{ref-StehmanKeyissuesrigorous2019}{}%
Stehman, S. V., and G. M. Foody. 2019. Key issues in rigorous accuracy
assessment of land cover products. Remote Sensing of Environment
231:111199.

\leavevmode\hypertarget{ref-sulla-menasheHierarchicalMappingAnnual2019}{}%
Sulla-Menashe, D., J. M. Gray, S. P. Abercrombie, and M. A. Friedl.
2019. Hierarchical mapping of annual global land cover 2001 to present:
{The MODIS Collection} 6 {Land Cover} product. Remote Sensing of
Environment 222:183--194.

\leavevmode\hypertarget{ref-Tongforgottenlanduse2020}{}%
Tong, X., M. Brandt, P. Hiernaux, S. Herrmann, L. V. Rasmussen, K.
Rasmussen, F. Tian, T. Tagesson, W. Zhang, and R. Fensholt. 2020. The
forgotten land use class: {Mapping} of fallow fields across the {Sahel}
using {Sentinel}-2. Remote Sensing of Environment 239:111598.

\leavevmode\hypertarget{ref-tuiaSurveyActiveLearning2011}{}%
Tuia, D., M. Volpi, L. Copa, M. Kanevski, and J. Munoz-Mari. 2011. A
{Survey} of {Active Learning Algorithms} for {Supervised Remote Sensing
Image Classification}. IEEE Journal of Selected Topics in Signal
Processing 5:606--617.

\leavevmode\hypertarget{ref-vanvlietThereContinuingRationale2013}{}%
Van Vliet, N., O. Mertz, T. Birch-Thomsen, and B. Schmook. 2013. Is
{There} a {Continuing Rationale} for {Swidden Cultivation} in the 21st
{Century}? Human Ecology 41:1--5.

\leavevmode\hypertarget{ref-visvalingamLineGeneralisationRepeated1993}{}%
Visvalingam, M., and J. D. Whyatt. 1993. Line generalisation by repeated
elimination of points. The Cartographic Journal 30:46--51.

\leavevmode\hypertarget{ref-VonBraunSmallscalefarmersliberalized2004}{}%
Von Braun, J. 2004. Small-scale farmers in a liberalized trade
environment. Page 21.

\leavevmode\hypertarget{ref-waldnerDeepLearningEdge2020}{}%
Waldner, F., and F. I. Diakogiannis. 2020. Deep learning on edge:
{Extracting} field boundaries from satellite images with a convolutional
neural network. Remote Sensing of Environment 245:111741.

\leavevmode\hypertarget{ref-WaldnerConflationexpertcrowd2019}{}%
Waldner, F., A. Schucknecht, M. Lesiv, J. Gallego, L. See, A.
Pérez-Hoyos, R. d'Andrimont, T. de Maet, J. C. L. Bayas, S. Fritz, O.
Leo, H. Kerdiles, M. Díez, K. Van Tricht, S. Gilliams, A. Shelestov, M.
Lavreniuk, M. Simões, R. Ferraz, B. Bellón, A. Bégué, G. Hazeu, V.
Stonacek, J. Kolomaznik, J. Misurec, S. R. Verón, D. de Abelleyra, D.
Plotnikov, L. Mingyong, M. Singha, P. Patil, M. Zhang, and P. Defourny.
2019. Conflation of expert and crowd reference data to validate global
binary thematic maps. Remote Sensing of Environment 221:235--246.

\leavevmode\hypertarget{ref-WilsonRemotelySensedHighResolution2016}{}%
Wilson, A. M., and W. Jetz. 2016. Remotely {Sensed High}-{Resolution
Global Cloud Dynamics} for {Predicting Ecosystem} and {Biodiversity
Distributions}. PLOS Biology 14:e1002415.

\leavevmode\hypertarget{ref-WulderglobalLandsatarchive2016}{}%
Wulder, M. A., J. C. White, T. R. Loveland, C. E. Woodcock, A. S.
Belward, W. B. Cohen, E. A. Fosnight, J. Shaw, J. G. Masek, and D. P.
Roy. 2016. The global {Landsat} archive: {Status}, consolidation, and
direction. Remote Sensing of Environment 185:271--283.

\leavevmode\hypertarget{ref-XiongNominal30mCropland2017}{}%
Xiong, J., P. S. Thenkabail, J. C. Tilton, M. K. Gumma, P. Teluguntla,
A. Oliphant, R. G. Congalton, K. Yadav, and N. Gorelick. 2017. Nominal
30-m {Cropland Extent Map} of {Continental Africa} by {Integrating
Pixel}-{Based} and {Object}-{Based Algorithms Using Sentinel}-2 and
{Landsat}-8 {Data} on {Google Earth Engine}. Remote Sensing 9:1065.

\leavevmode\hypertarget{ref-yeReviewAccuracyAssessment2018}{}%
Ye, S., R. G. Pontius, and R. Rakshit. 2018. A review of accuracy
assessment for object-based image analysis: {From} per-pixel to
per-polygon approaches. ISPRS Journal of Photogrammetry and Remote
Sensing 141:137--147.

\leavevmode\hypertarget{ref-YizongChengMeanshiftmode1995a}{}%
Yizong Cheng. 1995. Mean shift, mode seeking, and clustering. IEEE
Transactions on Pattern Analysis and Machine Intelligence 17:790--799.

\leavevmode\hypertarget{ref-ZengHighlandcroplandexpansion2018}{}%
Zeng, Z., L. Estes, A. D. Ziegler, A. Chen, T. Searchinger, F. Hua, K.
Guan, A. Jintrawet, and E. F. Wood. 2018. Highland cropland expansion
and forest loss in {Southeast Asia} in the twenty-first century. Nature
Geoscience 11:556--562.

\leavevmode\hypertarget{ref-Zhangimagetransformcharacterize2002}{}%
Zhang, Y., B. Guindon, and J. Cihlar. 2002. An image transform to
characterize and compensate for spatial variations in thin cloud
contamination of {Landsat} images. Remote Sensing of Environment
82:173--187.

\leavevmode\hypertarget{ref-ZhuDeeplearningremote2017}{}%
Zhu, X. X., D. Tuia, L. Mou, G.-S. Xia, L. Zhang, F. Xu, and F.
Fraundorfer. 2017. Deep learning in remote sensing: A review. IEEE
Geoscience and Remote Sensing Magazine 5:8--36.

\leavevmode\hypertarget{ref-ZhuObjectbasedcloudcloud2012}{}%
Zhu, Z., and C. E. Woodcock. 2012. Object-based cloud and cloud shadow
detection in {Landsat} imagery. Remote Sensing of Environment
118:83--94.

\end{CSLReferences}

\eleft

\clearpage

\newpage

\end{document}