### 2021-10-07 inst/manuscript/figures.Rmd
Revisions for Frontiers in AI round 1 reviews

---
title: Figures
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    math: katex
    toc_depth: 4
    toc: yes
    css: css/vignette.css
---

```{r setup, include=FALSE, cache=FALSE, message = FALSE}
library(activemapper)
library(raster)
library(sf)
library(stars)
library(dplyr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(RColorBrewer)
library(here)
library(glue)
library(aws.s3)
library(RStoolbox)
# library(gridExtra)
library(ggpmisc)
library(patchwork)

pth <- function(x) file.path(here("inst/manuscript"), x)
```

Creation of figures for main manuscript and the supplementary information.

## Base Datasets

The preparation of datasets used here and which install with the `activemapper` package, is shown in `data-raw/data_[from|to]_instances.R`. These are loaded in here.

```{r, eval=FALSE}
res <- 0.005
res_supergrid <- 1
num_supergrid <- res_supergrid / res

mgridr <- raster(
  system.file("extdata/ghana_grid.tif", package = "activemapper")
)
mgrid <- readr::read_csv(
  system.file("extdata", "ghana_grid.csv", package = "activemapper")
) %>% dplyr::select(-fwts)
africa <- st_read(
  system.file("extdata/africa.geojson", package = "activemapper")
)
africa_cont <- africa %>% st_union %>% st_sf
ghana <- africa %>% filter(name == "Ghana") %>% dplyr::select(geometry)

degrees <- st_read(
  system.file("extdata/degrees.geojson", package = "activemapper")
)
tiles <- st_read(
  system.file("extdata/ghana_tiles.geojson", package = "activemapper")
) 
aois <- st_read(
  system.file("extdata/aois.geojson", package = "activemapper")
)
aez <- st_read(
  system.file("extdata/ghana_aez.geojson", package = "activemapper")
)

# ggplot(data = tiles) + 
#   geom_sf(aes(fill = as.factor(aoi1)), col = "transparent")

# database results
load(system.file("extdata/", "instance_tbls.rda", package = "activemapper"))

# Ghana data
ghana_aois <- tiles %>% pull(aoi1) %>% unique
ghana_degrees <- tiles %>% pull(aoi2) %>% unique
```

## Methods

### Figure 1: System architecture
Created in Keynote

### Figure 2: Labeller interface
Taken as a screenshot.

### Figure S1: Grid reference system

```{r, eval = FALSE}
ind <- 8  # index of selected AOI for plotting (from ghana_aois)
xylims <- st_coordinates(ghana)
maprange <-  c(range(xylims[, 1]), range(xylims[, 2]))

p1 <- ggplot() + 
  geom_sf(data = africa, fill = "gray90", color = "grey50", size = 0.2) +
  geom_sf(data = ghana, fill = "grey85", col = "transparent", size = 0.3) +
  geom_sf(data = degrees %>% filter(grid %in% ghana_degrees),
          fill = "transparent", col = "grey70", size = 0.4, lty = 3) +
  geom_sf(data = aois, fill = "transparent", col = "black", size = 0.4) +
  geom_sf(data = aois %>% filter(aois == ind), col = "black", fill = "gray70", 
          size = 0.4) +
  geom_sf_text(data = aois, aes(label = aois)) + 
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) +
  theme_void()

deg_sel <- tiles %>% as_tibble %>% filter(aoi1 %in% 4:12) %>% 
  distinct(aoi1, aoi2) %>% pull()
deg_sel <- c(1349, 1350, 1351, 1414, 1415, 1416, 1478, 1479, 1480)

xylims <- degrees %>% filter(grid %in% deg_sel) %>% st_coordinates(.)
maprange <-  c(range(xylims[, 1]), range(xylims[, 2]))

tile_sel <- tiles %>% filter(aoi1 == ind)  # selected tile
p2 <- ggplot(ghana) + geom_sf(fill = "grey85", col = "grey") +
  geom_sf(data = degrees %>% filter(grid %in% ghana_degrees),
          fill = "transparent", col = "grey70", size = 0.4, lty = 3) +
  geom_sf(data = aois, fill = "transparent", size = 0.3) +
  geom_sf(data = tile_sel, fill = "gray70", size = 0.3) +
  geom_sf(data = tile_sel %>% slice(190), fill = "red3", size = 0.3) +
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) +
  theme_void()
# p2

gpols <- crop(mgridr, tile_sel[190, ]) %>% rasterToPolygons(.) %>%
  st_as_sf %>% st_sf(crs = 4326)
p3 <- ggplot(tile_sel %>% slice(169:171, 189:191, 209:211)) +
  geom_sf(fill = "grey80", size = 0.3) +
  geom_sf(data = gpols, fill = "red3", size = 0.3) +
  theme_void()
# p3

gp <- cowplot::plot_grid(p1, p2, p3, nrow = 1, scale = c(1, 1, 0.7), 
                         labels = LETTERS[1:3], label_x = 0.05)
ggsave(pth("figures/reference_system.png"), width = 7, height = 3,
       units = "in", dpi = 300)
```

### Figure S2: Bayesian Risk

A plot of the Bayesian risk curves associated with different levels of consensus, for both a cropland and non-cropland consensus label.
```{r, eval=FALSE}
# Heat map function
consensus <- seq(0, 1, 0.05)
risk <- function(r1, r2) {
  r1 * (1 - r2) + (1 - r1) * r2
}

risk_tbl <- tibble(
  Consensus = consensus, `Non-field` = risk(consensus, 0), 
  Field = risk(consensus, 1)
) %>% pivot_longer(cols = contains("field"), names_to = "Label", 
                   values_to = "Risk")
p <- ggplot(risk_tbl) + geom_line(aes(x = Consensus, y = Risk, color = Label)) +   scale_color_manual(values = c("blue", "red")) + 
  theme_linedraw()

ggsave(pth("figures/si_label_risk.png"), width = 5, 
       height = 4, units = "in", dpi = 300, bg = "transparent")
```

### Figure S3: Segmentation workflow 
Created in Keynote.

### Figure S4: Map reference sample
```{r, eval = FALSE}
ref_sample <- readr::read_csv(
  system.file("extdata", "map_reference_sample.csv", package = "activemapper")
)
ref_sample <- ref_sample %>% st_as_sf(coords = c("x", "y"), crs = 4326)

xylims <- st_coordinates(ghana)
maprange <-  c(range(xylims[, 1]), range(xylims[, 2]))

p <- ggplot(africa) + 
  geom_sf(fill = "gray97", color = "grey", size = 0.25) +
  geom_sf(data = aois, fill = "grey85", col = "black", size = 0.3) +
  geom_sf(data = ref_sample, aes(color = factor(class)), size = 0.1) +
  scale_color_manual(values = c("red", "blue"), 
                     labels = c("Non-cropland", "Cropland"), name = "") + 
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) +
  guides(colour = guide_legend(override.aes = list(size = 1, fill = NA))) + 
  theme_void() + 
  theme(legend.key = element_rect(fill = "grey85", color = "transparent"))

ggsave(pth("figures/si_map_reference_sample.png"), width = 4, 
       height = 4, units = "in", dpi = 300, bg = "transparent")
```

### Figure S5: AOI Zones

```{r, eval=FALSE}
# zone maps
zone_grps <- list("aoi" = list(1:3, 4:9, c(10:11, 13, 14, 16), c(12, 15)),
                  "aez" = list(1, 2:4, 5, 6:7))
aoi_zones <- aois %>% 
  mutate(zone = case_when(
    aois %in% zone_grps$aoi[[1]] ~ "Zone 1", 
    aois %in% zone_grps$aoi[[2]] ~ "Zone 2", 
    aois %in% zone_grps$aoi[[3]] ~ "Zone 3", 
    aois %in% zone_grps$aoi[[4]] ~ "Zone 4")
  ) %>%
  group_by(zone) %>% 
  summarize(area = sum(area))

rmapshaper::ms_simplify(aez) %>% plot()
aez %>% st_buffer(dist = 0.01) %>% st_union(by_feature = TRUE) %>% 
  st_buffer(dist = -0.01) %>% 
  mutate(zone = case_when(
    ID %in% zone_grps$aez[[1]] ~ "Zone 1", 
    ID %in% zone_grps$aez[[2]] ~ "Zone 2", 
    ID %in% zone_grps$aez[[3]] ~ "Zone 3", 
    ID %in% zone_grps$aez[[4]] ~ "Zone 4")
  ) %>%
  group_by(zone) %>% 
  summarize(area = sum(AREA)) %>% plot()

aez_zones <- rmapshaper::ms_simplify(aez, keep = 1) %>% 
  mutate(zone = case_when(
    ID %in% zone_grps$aez[[1]] ~ "Zone 1", 
    ID %in% zone_grps$aez[[2]] ~ "Zone 2", 
    ID %in% zone_grps$aez[[3]] ~ "Zone 3", 
    ID %in% zone_grps$aez[[4]] ~ "Zone 4")
  ) %>%
  group_by(zone) %>% 
  summarize(area = sum(AREA)) 
# plot(aez_zones["area"])
# plot(aez["AREA"])

m <- ggplot(aois) + geom_sf(lwd = 0.1) +
  geom_sf(data = aoi_zones, lwd = 0.5, fill = "transparent") + 
  geom_sf_text(data = st_centroid(aoi_zones), aes(label = zone), size = 4) + 
  theme_void() + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))

m2 <- ggplot(aez %>% filter(ID != 8)) + geom_sf(lwd = 0.1) +
  geom_sf(data = aez_zones, lwd = 0.5, fill = "transparent") + 
  geom_sf_text(data = st_centroid(aez) %>% filter(ID != 8) %>% arrange(ID),
               aes(label = ID), 
               nudge_x = c(0.75, 0, 0, -1.25, -1, -1.25, 0),
               nudge_y = c(0.25, 0, 0, 0.7, 0.25, 1, 0),
               size = 3, col = "grey60") + 
  geom_sf_text(data = st_centroid(aez_zones), aes(label = zone), size = 4) + 
  theme_void() + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
ggsave(m + m2 + plot_annotation(tag_levels = 'A'), 
       filename = pth("figures/si_aoi_zones.png"), width = 3.5 * 2, 
       height = 4.5, units = "in", dpi = 300)
```

## Results 
### Figure 3: Composite quality
```{r, eval=FALSE}
data("image_quality")

# plot parameters
xylims <- st_coordinates(ghana)
maprange <-  c(range(xylims[, 1]), range(xylims[, 2]))
pal <- colorRampPalette(rev(brewer.pal(9, "RdBu")))
# sc <- scale_colour_gradientn(colours = myPalette(100), limits=c(1, 8))

leglabs <- seq(0.4, 1, 0.1)
p1 <- ggplot(africa) + geom_sf(fill = "gray97", color = "grey", size = 0.25) +
  geom_sf(data = tiles, fill = "grey85", col = "grey", size = 0.3) +
  geom_sf(data = ghana, fill = "transparent", col = "grey", size = 0.3) +
  geom_sf(data = image_quality$tile, 
          mapping = aes(fill = Score, color = Score)) +
  facet_wrap(vars(season), ncol = 2) +
  scale_fill_gradientn(colors = pal(7), 
                       breaks = seq(0.4, 1, 0.1), limits = c(0.4, 1),
                       labels = leglabs) + 
  scale_color_gradientn(colors = pal(7), 
                        breaks = seq(0.4, 1, 0.1), limits = c(0.4, 1), 
                        labels = leglabs) + 
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) + 
  theme_linedraw() + 
  theme(legend.position = "bottom", 
        strip.background = element_rect(fill = "white"), 
        strip.text = element_text(colour = "black"), 
        panel.grid = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank())

p2 <- ggplot(as_tibble(image_quality$tile)) +
  geom_histogram(aes(x = Score), fill = "blue", col = "grey", #bins = 10, 
                 binwidth = 0.05, boundary = 0.1) + 
  scale_x_continuous(breaks = leglabs, labels = leglabs, expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 30)) +
  facet_wrap(vars(season), ncol = 2) + 
  ylab("Count") + 
  theme_linedraw() + 
  theme(strip.background = element_rect(fill = "white"), 
        strip.text = element_text(colour = "black"))

g <- cowplot::plot_grid(p1, p2, nrow = 2, align = "hv", axis = "l",
                        labels = LETTERS[1:2], label_x = 0, 
                        label_y = c(0.97, 1),   
                        scale = c(1, 1), rel_heights = c(1, 0.7))
ggsave(pth("figures/figure3.png"), 
       width = 4, height = 6.5, units = "in", dpi = 300)

```

### Figure 4: Active learning gain
```{r, eval=FALSE}
# iteration metrics
data("iteration_metrics")

# Plots
V1 <- c("Accuracy", "AUC", "F1")

# prepare data
# remove labeller from AOI and select AOIs 1-16
aoinms <- c(paste0("labeller", 1:16), "All")
imetrics <- iteration_metrics %>% 
  filter(Metric %in% V1) %>% 
  filter(AOI %in% aoinms) %>% 
  mutate(AOI = gsub("labeller", "", AOI))

# select out final iteration for each AOI
mx_metrics <- imetrics %>% group_by(AOI) %>% 
  filter(Iteration == max(Iteration)) %>% filter(Metric %in% V1)

# plot
cols <- c(rep("grey", length(unique(imetrics$AOI)) -1), "black")
lsize <- c(rep(0.3, length(unique(imetrics$AOI)) -1), 1)
p <- ggplot(imetrics) + 
  geom_line(aes(Iteration, Score, color = AOI, size = AOI)) +
  geom_point(data = mx_metrics, aes(Iteration, Score, color = AOI), 
             size = 1.2) +
  ggrepel::geom_text_repel(
    data = mx_metrics %>% filter(AOI != "All"), 
    aes(label = AOI, x = Iteration, y = Score), col = "grey50", 
    segment.color = "grey60", point.padding = 0.1, min.segment.length = 0.2,
    segment.size = 0.2, size = 3   
  ) +
  scale_color_manual(values = cols, guide = FALSE) +
  scale_size_manual(values = lsize, guide = FALSE) + 
  scale_x_continuous(expand = c(0.01, 0.01)) + 
  scale_y_continuous(breaks = seq(0.2, 1, 0.1), expand = c(0.01, 0.01), 
                     limits = c(0.2, 1)) + 
  facet_wrap(vars(Metric), ncol = 3) + 
  theme_linedraw() + 
  theme(strip.background = element_rect(fill = "white"), 
        strip.text = element_text(colour = "black"))

ggsave(pth("figures/figure4.png"), 
       width = 7, height = 4, units = "in", dpi = 300)
```

### Figure 5: The impact of training data error
```{r, eval=FALSE}
data("consensus_high_low")

V1 <- c("Accuracy", "AUC", "F1")

# calculate mean across AOIs and join. Convert to factor and reorder for
# plotting
consensus_high_low_mu <- consensus_high_low %>% 
  group_by(Strategy, Metric) %>% 
  summarize(Score = mean(Score)) %>% 
  ungroup() %>% 
  mutate(AOI = "Mean") %>% 
  bind_rows(consensus_high_low, .) %>% 
  filter(Metric %in% V1) %>% 
  mutate(AOI = factor(AOI)) %>% 
  mutate(AOI = forcats::fct_reorder(AOI, ord))

cols <- c("grey50", "grey60", "grey70", "grey80", "black")
sizes <- c(rep(1.5, 4), 2)
shapes <- c(3, 4, 8, 10, 16)
p <- ggplot(consensus_high_low_mu) + 
  geom_point(aes(x = Strategy, y = Score, color = AOI, size = AOI, 
                 shape = AOI)) + 
  ylim(0.4, 1) + xlab("Labelling strategy") + 
  scale_color_manual(values = cols) +
  scale_size_manual(values = sizes) + 
  scale_shape_manual(values = shapes) + 
  facet_wrap(~Metric) + 
  theme_linedraw() + 
  theme(legend.position = "bottom", 
        strip.background = element_rect(fill = "white"), 
        strip.text = element_text(colour = "black"))
ggsave(pth("figures/figure5.png"), height = 4, width = 6, dpi = 300)
```

### Table 2: Categorical map accuracy
Place map accuracy tables and zone report into single figure, for AOI zones.

```{r, eval=FALSE}
data("accuracy_reports")

# RF accuracy
rfreports <-  lapply(accuracy_reports$zones$aoi, function(x) x$rf)
rfreports[[length(rfreports) + 1]] <- accuracy_reports$ghana$rf
rf_tabs <- lapply(1:5, function(i) {  # i <- 2
  m <- rfreports[[i]]
  d <- data.frame(m, check.names = FALSE)
  rbind(d, rep("", ncol(d)))
}) %>% do.call(rbind, .)
zvec <- rep("", nrow(rf_tabs))
znames <- paste(c(paste("Zone", 1:4), "Ghana"), "  ")
zvec[c(3, seq(8, length(zvec), by = 5))] <- znames
rf_tabs <- cbind(zvec, rf_tabs)
colnames(rf_tabs) <- c("", colnames(rfreports$North))

# rf_tbl <- tibble(x = 0, y = 10, dat = list(rf_tabs))

# segmentation accuracy
segreports <-  lapply(accuracy_reports$zones$aoi, function(x) x$seg)
segreports[[length(segreports) + 1]] <- accuracy_reports$ghana$seg
seg_tabs <- lapply(1:5, function(x) {  # x <- 1
  m <- segreports[[x]]
  d <- data.frame(m, check.names = FALSE)
  rbind(d, rep("", ncol(d)))
}) %>% do.call(rbind, .)
seg_tabs <- cbind(zvec, seg_tabs)
colnames(seg_tabs) <- c("", colnames(rfreports$North))#rep("", ncol(seg_tabs))
# seg_tbl <- tibble(x = 0, y = 10, dat = list(seg_tabs))

all_tbl <- tibble(x = 0, y = 10, dat = list(rbind(rf_tabs, seg_tabs)))

# Assemble plots with maps
# Table theme
tt <- ttheme_gtminimal(
  base_size = 10, 
  padding = unit(c(0.7, 0.3), "char"),
  core = list(fg_params = list(hjust = 1, x = 1)), 
  colhead = list(fg_params = list(hjust = 1, x = 1))
)

t1 <- ggplot(data.frame(x = c(0, 10), y = c(0, 10))) + 
  geom_point(aes(x = x, y = y), shape = "") +  
  geom_table(data = all_tbl, aes(x = x, y = y, label = dat), table.theme = tt) + 
  # ggtitle("Per-pixel classification") + 
  annotate(geom = "text", x = -0.2, y = 7, #y = 1, yend = 5, 
           label = "Per-pixel classification", angle = 90) +
  annotate(geom = "text", x = -0.2, y = 2.5, label = "Segmentation", 
           angle = 90, hjust = 1) +
  annotate(geom = "segment", x = -0.2, xend = 10, y = 9.77, yend = 9.77) + 
  annotate(geom = "segment", x = -0.2, xend = 10, y = 10.02, yend = 10.02) + 
  annotate(geom = "segment", x = -0.2, xend = 10, y = 4.7, yend = 4.7, 
           size = 0.3) + 
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 6), 
        plot.margin = unit(c(0, 0, 0, 0), "cm"))

# gp <- t1 + m
ggsave(t1, filename = pth("figures/table2.png"), width = 7, height = 7.9, 
       units = "in", dpi = 300)
```

### Figure 6: Cropland maps
```{r, eval=FALSE}
# segment files
segment_files <- dir(here("external/data/results/segments/merged"),
                     pattern = "boundarymerge", full.names = TRUE)
iid <- as.numeric(gsub("[[:alpha:]]|_|\\.", "", basename(segment_files)))
segment_files <- tibble(file = segment_files, order = iid) %>% arrange(iid)
segment_files <- segment_files %>% filter(!grepl("aoi3_*.*merge.geojson", file))

# fdensity <- stars::read_stars(
#   system.file("extdata", "field_density005.tif", package = "activemapper")
# )

# load in Ghana cropland fractions, mask out tiles over ocean, convert to stars
# percent
cropfrac <- raster(
  system.file("extdata/cropfrac_005.tif", package = "activemapper")
)
cropfrac <- africa %>% 
  filter(name %in% c("Ghana", "Togo", "Cte d'Ivoire", "Burkina Faso")) %>% 
  mask(cropfrac, .)
cropfrac <- st_as_stars(cropfrac * 100)

# individual fractions
tile_list <- c(486794, 539730, 591243, 616490)
tile_ss <- tiles %>% filter(tile %in% tile_list)

data("tile_key")
s3root <- "/vsis3/activemapper"
prob_tiles <- tile_key %>% filter(tile %in% tile_list) %>% 
  mutate(impath = glue("{s3root}/classified-images/{aoi}_whole/", 
                       "image_c{col}_r{row}.tif"))

# OS images
osimgs <- get_bucket_df(
  bucket = "activemapper", prefix = "planet/composite_sr_buf_fix/OS/", max = Inf
) %>% as_tibble %>% dplyr::select(Key)
impath <- osimgs %>% 
  filter(grepl(paste0(tile_ss$tile, collapse = "|"), Key)) %>% 
  mutate(pth = glue("{s3root}/{Key}")) %>% pull(pth)

# example tiles
# aoi 1: tile 486794
# aoi 5: tile 539730
# aoi 11: 
# aoi 15:

# read in field segments
flds_1 <- read_sf(segment_files$file[1]) %>% st_make_valid()
flds_5 <- read_sf(segment_files$file[5]) %>% st_make_valid()
flds_11 <- read_sf(segment_files$file[11]) %>% st_make_valid()
flds_15 <- read_sf(segment_files$file[15]) %>% st_make_valid()

# ggplot(aois %>% filter(aois == 1)) + geom_sf() + 
#   geom_sf(data = tile_ss)
flds_tile1 <- flds_1 %>% filter(st_intersects(., tile_ss[1, ], sparse = FALSE))
flds_tile2 <- flds_5 %>% filter(st_intersects(., tile_ss[2, ], sparse = FALSE))
flds_tile3 <- flds_11 %>% filter(st_intersects(., tile_ss[3, ], sparse = FALSE))
flds_tile4 <- flds_15 %>% filter(st_intersects(., tile_ss[4, ], sparse = FALSE))

# process inset maps
# img <- read_stars(impath)
# mxval <- st_apply(img, 3, max) %>% pull(max)
# imgngb <- st_rgb(img[, , , 4:2], 3, maxColorValue = mxval[4:2])
vpal <- viridis::viridis(20)
# inset_cols <- c("red", "orange", "pink", "brown")
inset_cols <- rep("red", 4)
fld_tiles <- list(flds_tile1, flds_tile2, flds_tile3, flds_tile4)
isets <- lapply(1:4, function(x) {  # x <- 1
  imgr <- brick(impath[x])[[2:4]]
  probr <- stars::read_stars(prob_tiles$impath[x])
  xylims <- st_bbox(imgr)
  iset <- ggplot() + 
    ggRGB(imgr, maxpixels = 50000, stretch = "lin", ggLayer = TRUE) +
    geom_sf(data = fld_tiles[[x]], fill = "transparent", color = "yellow", 
            size = 0.2) + 
    geom_sf(data = tile_ss[x, ], color = inset_cols[x], fill = "transparent") + 
    scale_x_continuous(expand = c(0, 0)) + 
    scale_y_continuous(expand = c(0, 0)) + 
    coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + 
    theme_void()
  iset2 <- ggplot() + 
    geom_stars(data = probr) +
    scale_fill_gradientn(name = "", colours = vpal, limits = c(0, 100)) +
    geom_sf(data = tile_ss[x, ],  color = inset_cols[x], fill = "transparent") +
    scale_x_continuous(expand = c(0, 0)) + 
    scale_y_continuous(expand = c(0, 0)) + 
    coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + 
    theme_void() + theme(legend.position = "none")
  cowplot::plot_grid(iset2, iset, ncol = 1, align = "hv")
})
# isets[[2]]

# Main map
xylims <- st_coordinates(ghana)
# twk <- diff(range(xylims[, 1])) * 0.05
maprange <-  c(range(xylims[, 1]) + c(-1.2, 1.2), 
               range(xylims[, 2]))

# create lines to point to insets
tile_ss_xy <- st_coordinates(st_centroid(tile_ss))
lns <- st_sfc(
  list(
    st_linestring(
      rbind(tile_ss_xy[1, ], c(maprange[1] * 0.9, maprange[4] * 0.885))
    ), 
    st_linestring(
      rbind(tile_ss_xy[2, ], c(maprange[2] * 0.9, maprange[4] * 0.8517))
    ),
    st_linestring(
      rbind(tile_ss_xy[3, ], c(maprange[1] * 0.9, maprange[4] * 0.619))
    ),
    st_linestring(
      rbind(
        tile_ss_xy[4, ], c(maprange[2] * 0.9, maprange[4] * 0.5681))
    )
  ), crs = 4326
)

# cropland fractions
p <- ggplot(africa) + 
  geom_sf(fill = "gray60", color = "grey", size = 0.25) +  
  geom_stars(data = cropfrac) +
  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
  geom_sf(data = tile_ss, fill = inset_cols, col = inset_cols, size = 0.1) +
  geom_sf(data = lns, col = inset_cols) + 
  geom_sf(data = tile_ss, fill = "red", col = "red") +
  scale_fill_viridis_c(name = "%", na.value = "transparent") +
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) + 
  ggsn::north(symbol = 16, scale = 1.2, x.min = 0.5, x.max = 1, 
              y.min = 10.5, y.max = 11) +
  ggsn::scalebar(ghana, dist = 100, dist_unit = "km", location = "bottomleft", 
                 transform = TRUE, model = "WGS84", st.size = 3, st.dist = 0.01, 
                 anchor = c(x = -0.7, y = 4.8), border.size = 0.1) + 
  theme_void() + 
  theme(legend.position = c(0.675, 0.1), legend.direction = "horizontal",
        legend.key.height = unit(0.3, "cm"),
        legend.key.width = unit(0.67, "cm"),
        legend.text = element_text(colour = "black"),
        legend.text.align = 0.5,
        legend.title.align = 1,
        legend.spacing.y = unit(0.05, "cm"),
        legend.title = element_text(
          vjust = 1, size = 11, colour = "black"
        )
  )

# with insets added
wh <- c(0.4, 0.4)
lx <- -0.1
rx <- 0.7
inset_map <- ggdraw() +
  draw_plot(p) +
  draw_plot(isets[[1]], x = lx, y = 0.6, width = wh[1], height = wh[2]) + 
  draw_plot(isets[[2]], x = rx, y = 0.51, width = wh[1], height = wh[2]) + 
  draw_plot(isets[[3]], x = lx, y = 0.17, width = wh[1], height = wh[2]) + 
  draw_plot(isets[[4]], x = rx, y = 0.07, width = wh[1], height = wh[2])
# inset_map

ggsave(filename = pth("figures/figure6.png"), 
       plot = inset_map, bg = "transparent", width = diff(maprange[1:2]), 
       height = diff(maprange[3:4]), units = "in", dpi = 300)

```

### Figure S6: PlanetScope image density
The density of imagery available for download from the image catalog.

```{r imagedensity, eval=FALSE}
img_density <- st_read(
  system.file("extdata/img_catalog_count_201712_201902.geojson", 
              package = "activemapper")
)

tile_count <- img_density %>% mutate(season = case_when(
  year == 2018 & month %in% 5:9 ~ "GS",
  year == 2018 & month %in% 12 ~ "OS", 
  year == 2019 & month %in% 1:2 ~ "OS", 
  TRUE ~ NA_character_)
) %>% filter(season %in% c("GS", "OS")) %>% 
  as_tibble() %>%
  group_by(tile, season) %>%
  summarize(Count = sum(count))

gs_count <- tiles %>% dplyr::select(tile) %>% 
  left_join(., tile_count %>% filter(season == "GS"))
os_count <- tiles %>% dplyr::select(tile) %>% 
  left_join(., tile_count %>% filter(season == "OS")) # %>%
  # fasterize::fasterize(., field = "count")

pal <- colorRampPalette(rev(RColorBrewer::brewer.pal(9, "RdBu")))
zlim <- range(c(gs_count$Count, os_count$Count))
p <- ggplot() + 
  geom_sf(data = gs_count, aes(fill = Count), col = "transparent") + 
  geom_sf(data = aois, fill = "transparent", size = 0.2) + 
  scale_fill_gradientn(colours = pal(20), limits = zlim) +
  ggtitle("May - Sept 2018") + 
  theme_void() + 
  theme(legend.position = "none", 
        title = element_text(size = 9))

p2 <- ggplot() + 
  geom_sf(data = os_count, aes(fill = Count), col = "transparent") + 
  scale_fill_gradientn(colours = pal(20), limits = zlim) +
  geom_sf(data = aois, fill = "transparent", size = 0.2) + 
  ggtitle("Dec 2018 - Feb 2019") +
  theme_void() + 
  theme(title = element_text(size = 9),
        legend.text = element_text(size = 9))

op <- p + p2
ggsave(op, bg = "transparent",
       filename = pth("figures/si_planet_coverage.png"), 
       width = 5, height = 3, dpi = 300, units = "in")

```

### Figure S7: Initial training site distribution
Figure showing distribution of initial 500 samples per cluster, as well as distribution of points per iteration.  
```{r trainingsites, eval=FALSE}
data("train_val_sites")

qsites <- instance_tbls$kml_data %>% 
  filter(aoi %in% paste0("labeller", 1:16)) %>% 
  filter(kml_type == "Q") %>% select(name, x, y) %>% distinct() %>%
  mutate(Cluster = "Training reference") %>% 
  st_as_sf(coords = c("x", "y"), crs = 4326) %>% select(Cluster, name)

train_val_q <- rbind(train_val_sites$initial %>% select(-iteration), qsites)

# initial train cluster panel
xylims <- st_coordinates(ghana)
maprange <-  c(range(xylims[, 1]), range(xylims[, 2]))
pal <- RColorBrewer::brewer.pal(n = 9, "Spectral")[c(1, 2, 3, 4, 9)]

p1 <- ggplot(africa) + 
  geom_sf(fill = "gray97", color = "grey", size = 0.25) +
  geom_sf(data = aois, fill = "grey85", col = "black", size = 0.3) +
  geom_sf(data = train_val_q, aes(color = Cluster, fill = Cluster), 
          size = 0.05) +   
  geom_sf_text(data = st_centroid(aez) %>% filter(ID != 8) %>% arrange(ID),
               aes(label = ID), 
               nudge_x = c(0.75, 0, 0, -1.25, -1, -1.25, 0),
               nudge_y = c(0.25, 0, 0, 0.7, 0.25, 1, 0),
               size = 3, col = "grey60") + 
  geom_sf(data = aez, fill = "transparent", color = "gray60", size = 0.25) +
  scale_color_manual(values = pal) +
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) +
  guides(colour = guide_legend(override.aes = list(size = 1, fill = NA))) + 
  theme_void() + 
  theme(legend.key = element_rect(fill = "grey85", color = "transparent"))

pal <- c(RColorBrewer::brewer.pal(n = 4, "Spectral"), "white")
p2 <- ggplot(africa) + geom_sf(fill = "gray97", color = "grey", size = 0.25) +
  geom_sf(data = aois, fill = "grey85", col = "black", size = 0.3) +
  geom_sf(data = train_val_sites$active, aes(color = Type), size = 0.05) +
  geom_sf_text(data = st_centroid(aez) %>% filter(ID != 8) %>% arrange(ID),
               aes(label = ID), 
               nudge_x = c(0.75, 0, 0, -1.25, -1, -1.25, 0),
               nudge_y = c(0.25, 0, 0, 0.7, 0.25, 1, 0),
               size = 3, col = "grey60") + 
  geom_sf(data = aez, fill = "transparent", color = "gray60", size = 0.25) +
  scale_color_manual(values = pal) +
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) +
  guides(colour = guide_legend(override.aes = list(size = 1, fill = NA))) + 
  theme_void() + 
  theme(legend.key = element_rect(fill = "grey85", color = "transparent"))

# arrange and save
# cp_theme1 <- theme(legend.box.just = "left")
cp_theme1 <- theme(legend.margin = margin(0, 0, 0, 0, "cm"), 
                   legend.text.align = 0, legend.title.align = 0)
gp <- cowplot::plot_grid(p1 + cp_theme1, 
                         p2 + cp_theme1, 
                         nrow = 2,
                         align = "hv", #axis = "l",
                         labels = LETTERS[1:2], label_x = 0.05)
ggsave(pth("figures/si_training_validation_pts.png"), width = 4, 
       height = 6, units = "in", dpi = 300, bg = "transparent")

```

### Figure S8: Assignment count
```{r assignmentcount, eval=FALSE}
data("label_summary")

# combine with label_quality
label_quality2 <- left_join(
  label_summary$scores, label_summary$score_stats, by = "uuid"
) %>% select(uuid, score, N, Mean, StDev) %>% 
  mutate(uuid = substr(uuid, 1, 5)) %>% rename(ID = uuid)

# find Q counts to get IDs infrequent workers
toofewq <- label_quality2 %>% group_by(ID) %>% count() %>% filter(n < 10)
label_quality2 <- label_quality2 %>% filter(!ID %in% toofewq$ID)

# N sites by type
yl <- c(0, ceiling(max(label_summary$nlabels$n) * 1.01))
p <- label_summary$nlabels %>% filter(!ID %in% toofewq$ID) %>% 
  group_by(ID, kml_type) %>% 
  ggplot() + 
  geom_bar(aes(x = reorder(ID, -n), y = n, fill = kml_type), 
           position = "dodge", stat = "identity") +
  xlab("ID") + ylab("N") + 
  labs(fill = "Label Type") + 
  scale_fill_manual(labels = c("Training/validation", "Accuracy assessment"),
                    values = c("grey40", "grey70")) + 
  scale_y_continuous(breaks = seq(yl[1], yl[2], by = 200), 
                     expand = c(0, 0.01), limits = yl) + 
  theme_linedraw() + 
  # theme(legend.position = c(0.7, 0.85), axis.text = element_text(size = 11),
  theme(axis.text = element_text(size = 11), 
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

ggsave(pth("figures/si_training_n.png"), width = 5.5, 
       height = 4, units = "in", dpi = 300, bg = "transparent")
```

### Figure S9: Random versus active sampling
```{r, eval=FALSE}
data("iteration_metrics")

# Plots
V1 <- c("Accuracy", "AUC", "F1")

# prepare data
# remove labeller from AOI and select AOIs 1-16
aois <- c(paste0("labeller", c(1, 8, 15, "1r", "8r", "15r")))
imetrics <- iteration_metrics %>% 
  filter(Metric %in% V1 & AOI %in% aois) %>% 
  mutate(AOI = gsub("labeller", "", AOI)) %>% 
  mutate(type = case_when(
    AOI %in% c(1, 8, 15) ~ "active", 
    AOI %in% c("1r", "8r", "15r") ~ "random"
  )) %>% 
  mutate(AOI = gsub("r", "", AOI)) %>% 
  tidyr::pivot_wider(names_from = "type", values_from = "Score") %>% 
  mutate(delta = (active - random) / random * 100)

imetrics_all <- imetrics %>% filter(Iteration < 4) %>% 
  group_by(Iteration, Metric) %>% 
  summarize(active = mean(active), random = mean(random), 
            delta = mean(delta)) %>%
  ungroup() %>% mutate(AOI = "All") %>% 
  select(AOI, Iteration, Metric, active, random, delta)
imetrics <- bind_rows(imetrics, imetrics_all)

# select out final iteration for each AOI
mx_metrics <- imetrics %>% group_by(AOI) %>% 
  filter(Iteration == max(Iteration)) %>% filter(Metric %in% V1)

# plot
cols <- c(rep("grey50", length(unique(imetrics$AOI)) -1), "black")
lsize <- c(rep(0.3, length(unique(imetrics$AOI)) -1), 1)
p <- ggplot(imetrics) + 
  geom_hline(yintercept = 0, linetype = "dotted", col = "grey40") +
  geom_line(aes(Iteration, delta, color = AOI, size = AOI)) +
  geom_point(data = mx_metrics, aes(Iteration, delta, color = AOI), 
             size = 1.2) +
  ggrepel::geom_text_repel(
    data = mx_metrics %>% filter(AOI != "All"), 
    aes(label = AOI, x = Iteration, y = delta), col = "grey50", 
    segment.color = "grey60", point.padding = 0.2, min.segment.length = 0.2,
    segment.size = 0.2, size = 3, nudge_x = 0 
  ) +
  scale_color_manual(values = cols, guide = FALSE) +
  scale_size_manual(values = lsize, guide = FALSE) + 
  scale_x_continuous(expand = c(0.01, 0.01)) + 
  scale_y_continuous(expand = c(0.01, 0.01), breaks = seq(-8, 4, 1)) + 
  ylab("% difference") +                    
  facet_wrap(vars(Metric), ncol = 3) +
  theme_linedraw() + 
  theme(panel.grid.minor.x = element_blank(), 
        strip.background = element_rect(fill = "white"), 
        strip.text = element_text(colour = "black"))

ggsave(pth("figures/si_random_vs_active.png"), 
       width = 6, height = 4, units = "in", dpi = 300)

```

### Figure S10: Label quality
```{r labelquality, eval=FALSE}
# Plot data from Figure S7

# Q site accuracy
p <- ggplot(label_quality2) + 
  geom_boxplot(aes(x = reorder(ID, -score), y = score, fill = ID), 
               col = "grey50", varwidth = TRUE, show.legend = FALSE,
               outlier.size = 1) + 
  geom_point(aes(x = ID, y = Mean), shape = 4, size = 2) +
  scale_y_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1), 
                     expand = c(0, 0.01)) + 
  scale_fill_manual(values = rep("grey", length(unique(label_quality2$ID)))) + 
  xlab("ID") + ylab("Score") + 
  theme_linedraw() + 
  theme(axis.text = element_text(size = 11), 
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

ggsave(pth("figures/si_label_quality.png"), width = 4, 
       height = 4, units = "in", dpi = 300, bg = "transparent")
```

### Figures S11 and S12: Consensus label (Bayesian) risk

Plots of mean risk in space, divided into training and validation sites

```{r riskpoints, eval=FALSE}
data("label_risk")

# Training/validation points
label_risk_pts <- label_risk$pts %>% 
  st_as_sf(coords = c("x", "y"), crs = 4326)
# label_consensus %>% distinct(aoi)
# label_risk %>% filter(is.na(meanrisk))

# # prepare for histograms
# clusters <- paste0("labeller", c("", "trainsouth", "trainsouthnl", "trainsw"))
# new_cluster_nms <- paste0("Cl", c("1", "2a", "2b", "3"))
# 
# label_risk_tb <- as_tibble(label_risk_pts) %>% 
#   mutate(aoi = case_when(
#     aoi == clusters[1] ~ "-3",
#     aoi == clusters[2] ~ "-2",
#     aoi == clusters[3] ~ "-1",
#     aoi == clusters[4] ~ "0", 
#     TRUE ~ gsub("labeller", "", aoi) 
#   )) %>% 
#   mutate(ord = as.numeric(aoi)) %>% arrange(aoi) %>% 
#   mutate(aoi = case_when(
#     aoi == "-3" ~ new_cluster_nms[1], 
#     aoi == "-2" ~ new_cluster_nms[2],
#     aoi == "-1" ~ new_cluster_nms[3],
#     aoi %in% "0" ~ new_cluster_nms[4], 
#     TRUE ~ aoi 
#   )) %>% 
#   mutate(aoi = forcats::fct_reorder(aoi, ord))
# 
# # summarize
# label_risk_stats <- label_risk_tb %>% group_by(aoi, usage) %>% 
#   summarize(Risk = mean(Risk, na.rm = TRUE)) %>% 
#   ungroup()

xylims <- st_coordinates(ghana)
maprange <-  c(range(xylims[, 1]), range(xylims[, 2]))
pal <- colorRampPalette(rev(RColorBrewer::brewer.pal(9, "RdBu")))
zlim <- c(0, round(max(label_risk_pts$Risk, na.rm = TRUE), 1))

p <- ggplot(africa) + 
  geom_sf(fill = "gray97", color = "grey", size = 0.25) +
  geom_sf(data = aois, fill = "grey85", col = "black", size = 0.3) +
  geom_sf(data = label_risk_pts, aes(color = Risk, fill = Risk), 
          size = 0.1) +
  facet_wrap(~usage) + 
  scale_color_gradientn(colours = pal(20), limits = zlim) + 
  scale_fill_gradientn(colours = pal(20), limits = zlim) +
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) +
  guides(fill = guide_colorbar(nbin = 20), color = FALSE) +
  theme_void() + 
  theme(legend.key = element_rect(fill = "grey85", color = "transparent"))
ggsave(pth("figures/si_label_risk_map.png"), width = 7, 
       height = 5, units = "in", dpi = 300, bg = "transparent")
```

Risk histograms. 
```{r riskhists, eval=FALSE}
p <- ggplot(label_risk$pts) + 
  geom_histogram(aes(x = Risk), fill = "grey50", bins = 20) + 
  geom_vline(data = label_risk$stats, aes(xintercept = Risk), size = 0.5, 
             linetype = 1) + 
  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 150, 25), 
                     labels = c("", 25, "", 75, "", 125, ""), 
                     limits = c(0, 150)) + 
  ylab("Number of sites") + 
  facet_grid(aoi ~ usage) +
  theme_linedraw() + 
  theme(strip.background = element_rect(fill = "white"), 
        strip.text = element_text(size = 9, colour = "black"),  
        axis.text.x = element_text(size = 9, hjust = 1, vjust = 0.5, 
                                   angle = 90), 
        axis.text.y = element_text(size = 7))
ggsave(pth("figures/si_label_risk_hists.png"), width = 3, 
       height = 9, units = "in", dpi = 300, bg = "transparent")
```

```{r, eval=FALSE}
data("label_risk_metrics")

label_risk_metrics %>% 
  filter(Metric != "TSS") %>%  
  ggplot() + geom_point(aes(Risk, Score)) + 
  geom_smooth(aes(Risk, Score), method = "gam") + 
  facet_wrap(~Metric)

label_risk_metrics %>% select(aoi, Metric, Score) %>% 
  tidyr::pivot_wider(names_from = "Metric", values_from = "Score") %>% 
  ggplot() + geom_point(aes(Precision, Accuracy))

label_risk_metrics %>% select(aoi, Metric, Score) %>% 
  tidyr::pivot_wider(names_from = "Metric", values_from = "Score") %>% 
  ggplot() + geom_point(aes(Precision, F1))

label_risk_metrics %>% select(aoi, Metric, Score) %>% 
  tidyr::pivot_wider(names_from = "Metric", values_from = "Score") %>% 
  ggplot() + geom_point(aes(Recall, F1))

label_risk_metrics %>% select(aoi, Metric, Score) %>% 
  tidyr::pivot_wider(names_from = "Metric", values_from = "Score") %>% 
  ggplot() + geom_point(aes(Precision, Recall))

label_risk_metrics %>% select(aoi, Metric, Score) %>% 
  tidyr::pivot_wider(names_from = "Metric", values_from = "Score") %>% 
  ggplot() + geom_point(aes(Precision, FPR))
label_risk_metrics %>% select(aoi, Metric, Score) %>% 
  tidyr::pivot_wider(names_from = "Metric", values_from = "Score") %>% 
  ggplot() + geom_point(aes(Recall, FPR))

```

### Figure S13: Training data error impact on probability images
```{r, eval = FALSE}
probpal <- viridis::viridis
cprob <- raster(here::here("external/prob/image_c294_r532.tif"))
hprob <- raster(here::here("external/prob/image_c294_r532_high.tif"))
lprob <- raster(here::here("external/prob/image_c294_r532_low.tif"))

png(pth("figures/si_label_strategy_probs.png"), height = 2.5,
    width = 8, res = 300, units = "in", bg = "transparent")
par(mfrow = c(1, 3), mar = c(0, 0.5, 0, 1.75), oma = c(0, 0, 0, 2))
plot(as(extent(cprob), "SpatialPolygons"), lty = 0)
plot(cprob, lty = 0, add = TRUE, legend = FALSE, zlim = c(0, 100),
     col = probpal(20))
mtext("A", side = 3, line = -2.5, adj = -0.02, cex = 1)
plot(as(extent(cprob), "SpatialPolygons"), lty = 0, cex.main = 2)
plot(hprob, lty = 0, add = TRUE, legend = FALSE, zlim = c(0, 100),
     col = probpal(20))
mtext("B", side = 3, line = -2.5, adj = -0.02, cex = 1)
plot(as(extent(cprob), "SpatialPolygons"), lty = 0)
plot(lprob, lty = 0, add = TRUE, zlim = c(0, 100),
     legend.width = 1, legend.shrink = 0.5, col = probpal(20))
mtext("C", side = 3, line = -2.5, adj = -0.02, cex = 1)
dev.off()

```

### Table S4: AEZ accuracies

Grouped by four zones, calculated for per-pixel classifications only.

```{r, eval=FALSE}
data("accuracy_reports")

# RF accuracy
rfreports <- lapply(accuracy_reports$zones$aez, function(x) x$rf)
rf_tabs <- lapply(1:4, function(i) {  # i <- 2
  m <- rfreports[[i]]
  d <- data.frame(m, check.names = FALSE)
  rbind(d, rep("", ncol(d)))
}) %>% do.call(rbind, .)
zvec <- rep("", nrow(rf_tabs))
znames <- paste(paste("Zone", 1:4), "  ")
zvec[c(3, seq(8, length(zvec), by = 5))] <- znames
rf_tabs <- cbind(zvec, rf_tabs)
colnames(rf_tabs) <- c("", colnames(rfreports$Coast_Sav))

all_tbl <- tibble(x = 0, y = 10, dat = list(rf_tabs))

# Assemble plots with maps
# Table theme
tt <- ttheme_gtminimal(
  base_size = 10, 
  padding = unit(c(0.7, 0.3), "char"),
  core = list(fg_params = list(hjust = 1, x = 1)), 
  colhead = list(fg_params = list(hjust = 1, x = 1))
)

t1 <- ggplot(data.frame(x = c(0, 10), y = c(0, 10))) + 
  geom_point(aes(x = x, y = y), shape = "") +  
  geom_table(data = all_tbl, aes(x = x, y = y, label = dat), table.theme = tt) + 
  annotate(geom = "segment", x = 0, xend = 10, y = 9.5, yend = 9.5) + 
  annotate(geom = "segment", x = 0, xend = 10, y = 10.01, yend = 10.01) + 
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 6), 
        plot.margin = unit(c(0, 0, 0, 0), "cm"))

ggsave(t1, filename = pth("figures/si_aez_accuracies.png"), width = 7, 
       height = 3.5, units = "in", dpi = 300)
```

### Figures S14 and S15: Segmentation statistics (grid scale)

```{r segstats, eval=FALSE}
# data("top_label_data")
# data("field_validation_stats")
data("segment_quality_stats")
# data
datnms <- c("top_label_data", "field_validation_stats")#,
            # "tile_aoi_validation_stats")
for(i in datnms) {
  load(system.file(glue::glue("extdata/{i}.rda"), package = "activemapper"))
}

# extract distributions
human <- top_label_data$stats %>%
  mutate(mu_prop = NA, type = "Validation") %>%
  select(-contains("id"))
machine <- field_validation_stats %>%
  mutate(type = "Segmentation")
humachine <- bind_rows(human, machine) %>%
  mutate(aoi = as.numeric(gsub("labeller", "", aoi))) %>%
  arrange(aoi)

squal <- segment_quality_stats %>% 
  filter(scale == "grid" & aoi != "All") %>% 
  mutate(aoi = as.numeric(aoi))
# 
# # Summarize AOI-wise averages (field size)
# humachine_sum <- humachine %>% group_by(aoi, type) %>% 
#   summarize(Median = median(Mean), Mean = mean(Mean)) %>% 
#   ungroup() %>% 
#   tidyr::pivot_longer(cols = Median:Mean, names_to = "Stat")
# # field number
# humachine_sum_n <- humachine %>% group_by(aoi, type) %>% 
#   summarize(Median = median(n), Mean = mean(n)) %>% 
#   ungroup() %>% 
#   tidyr::pivot_longer(cols = Median:Mean, names_to = "Stat")

# plots
fcols <- c(Segmentation = rgb(0, 0, 1, alpha = 0.5),
          Validation = rgb(1, 0, 0, alpha = 0.5))
brks <- seq(0, 30, 5)
theme_fun <- function(x) {
  theme_linedraw() + 
  theme(legend.position = "bottom", legend.text = element_text(size = 11),
        strip.background = element_rect(fill = "white"), 
        strip.text = element_text(size = 11, colour = "black"),  
        axis.text.x = element_text(size = 11, hjust = 1, vjust = 0.5, 
                                   angle = 90), 
        axis.text.y = element_text(size = 11))
}

# Field size
p <- ggplot(humachine) + 
  geom_density(aes(x = Mean, y = ..count.., fill = type), col = "transparent", 
               linetype = "blank") +  
  geom_vline(data = squal %>% filter(measure == "area"), 
             aes(xintercept = value, color = type, linetype = Stat), 
             size = 0.5) +
  scale_x_continuous(expand = c(0.03, 0.03), breaks = brks, limits = c(0, 30)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_color_manual(name = "", values = c("blue", "red"), guide = FALSE) + 
  scale_fill_manual(name = "", values = fcols) +  
  # scale_linetype_manual(name = "", values = c("dashed", "solid")) +  
  scale_linetype("") +  
  facet_wrap(vars(aoi)) + 
  xlab("Field size (ha)") + ylab("Count") + 
  theme_fun()
ggsave(pth("figures/si_validation_stats_fha2.png"), 
       width = 7, height = 7, units = "in", dpi = 300)

# field number
brks <- seq(0, 20, 5)
p2 <- ggplot(humachine) + 
  geom_density(aes(x = n, y = ..count.., fill = type), col = "transparent", 
               linetype = "blank") +  
  geom_vline(data = squal %>% filter(measure == "n"), 
             aes(xintercept = value, color = type, linetype = Stat), 
             size = 0.5) +
  scale_x_continuous(expand = c(0.03, 0.03), breaks = brks, limits = c(0, 20)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 40)) +
  scale_color_manual(name = "", values = c("blue", "red"), guide = FALSE) + 
  scale_fill_manual(name = "", values = fcols) +  
  # scale_linetype_manual(name = "", values = c("dashed", "solid")) +  
  scale_linetype("") +  
  facet_wrap(vars(aoi)) + 
  xlab("Number of fields") + ylab("Count") + 
  theme_fun()
ggsave(pth("figures/si_validation_stats_fn2.png"), 
       width = 7, height = 7, units = "in", dpi = 300)

```

### Figure S16: Average field sizes

Tile scale field size and number, presented without adjustment

```{r, eval=FALSE}
fsizes_n_05 <- st_read(
  system.file("extdata/field_size_n_05.geojson", package = "activemapper")
) %>% mutate(
  muarea = ifelse(is.na(muarea), 0, muarea),
  nflds = ifelse(is.na(nflds), 0, nflds),
  aoi = as.character(aoi)
)

# # calculate bias factors for adjustment
# data("segment_quality_stats")
# segment_quality_stats %>% filter(measure == "n")
# biases <- segment_quality_stats %>%
#   filter(Stat == "Mean" & scale == "grid") %>% 
#   pivot_wider(names_from = type, values_from = value) %>% 
#   mutate(bias = Validation - Segmentation, 
#          biasfactor = Validation / Segmentation)
# biases <- as_tibble(tiles) %>% group_by(aoi1) %>% 
#   count() %>% ungroup() %>% 
#   rename(aoi = aoi1, ntiles = n) %>% 
#   mutate(aoi = as.character(aoi), ncells = ntiles * 100) %>% 
#   left_join(biases, .) %>% 
#   select(-scale, -Stat)
# # biases %>% filter(measure == "n")
# # tiles %>% slice(1) %>% raster::extract(mgridr, .)
# # st_area(aois) / 10000 / 100
# 
# # apply bias adjustments
# # field area first
# fsizes_n_05_adj <- biases %>% filter(measure == "area" & aoi != "All") %>% 
#   select(aoi, biasfactor) %>% 
#   rename(fbias = biasfactor) %>% 
#   left_join(fsizes_n_05, .) %>% 
#   mutate(muarea_adj = muarea * fbias) 
# 
# # then field count
# fsizes_n_05_adj <- biases %>% filter(measure == "n" & aoi != "All") %>% 
#   select(aoi, biasfactor, ntiles, ncells) %>% 
#   rename(nbias = biasfactor) %>% 
#   left_join(fsizes_n_05_adj, .) %>% 
#   mutate(nflds_adj = nflds * nbias) %>% 
#   select(aoi, tile, muarea, muarea_adj, fbias, nflds, nflds_adj, nbias, ntiles)
# 
# 
# # hist(fsizes_n_05$muarea)
# brks <- c(-1, 0.001, 1, 2, 4, 6, 10, 15, 25, 50)
# labs <- c("0", "<1", "1-2", "2-4", "4-6", "6-10", "10-15", "15-25", "25-50")
# labsv <- c(1:(length(labs)))
# fsizes_n_05_adj <- fsizes_n_05_adj %>% 
#   mutate(
#     Area = as.numeric(cut(muarea, breaks = brks, labels = labsv)), 
#     Area_adj = as.numeric(cut(muarea_adj, breaks = brks, labels = labsv))
#   )
# 
# nbrks <- c(-1, 0.001, 50, 100, 200, 300, 400, 500, 700, 1000)
# nlabs <- c("0", "<50", "50-100", "100-200", "200-300", "300-400", "400-500",
#            "500-700", "700-1000")
# nlabsv <- c(1:(length(labs)))
# fsizes_n_05_adj <- fsizes_n_05_adj %>% 
#   mutate(
#     N = as.numeric(cut(nflds, breaks = nbrks, labels = nlabsv)), 
#     N_adj = as.numeric(cut(nflds_adj, breaks = nbrks, labels = nlabsv))
#   )
# 
# # stars
# muarea <- st_rasterize(fsizes_n_05_adj["Area"])
# muarea_adj <- st_rasterize(fsizes_n_05_adj["Area_adj"])
# nflds <- st_rasterize(fsizes_n_05_adj["N"])
# nflds_adj <- st_rasterize(fsizes_n_05_adj["N_adj"])

brks <- c(-1, 0.001, 1, 2, 4, 6, 10, 15, 25, 50)
labs <- c("0", "<1", "1-2", "2-4", "4-6", "6-10", "10-15", "15-25", "25-50")
labsv <- c(1:(length(labs)))
fsizes_n_05 <- fsizes_n_05 %>% 
  mutate(Area = as.numeric(cut(muarea, breaks = brks, labels = labsv)))

# hist(fsizes_n_05$nflds)
nbrks <- c(-1, 0.001, 50, 100, 200, 300, 400, 500)
nlabs <- c("0", "<50", "50-100", "100-200", "200-300", "300-400", "400-500")
nlabsv <- c(1:(length(nlabs)))
fsizes_n_05 <- fsizes_n_05 %>% 
  mutate(N = as.numeric(cut(nflds, breaks = nbrks, labels = nlabsv)))

muarea <- st_rasterize(fsizes_n_05["Area"])
nflds <- st_rasterize(fsizes_n_05["N"])

thm <- function(x) {
  theme_void() + 
    theme(#legend.position = c(0.7, 0.07), 
          legend.direction = "vertical", 
          legend.key.height = unit(0.2, "cm"), 
          legend.text = element_text(colour = "black"),
          legend.title = element_text(size = 9, colour = "black"))
}

leg <- c(labs, "")
xylims <- st_bbox(ghana)
p <- ggplot(africa) + 
  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
  geom_stars(data = muarea, aes(x = x, y = y, fill = factor(Area))) +
  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
  scale_fill_viridis_d(name = "ha", na.value = "transparent", labels = leg) +
  coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + thm()
# padj <- ggplot(africa) + 
#   geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
#   geom_stars(data = muarea_adj, aes(x = x, y = y, fill = factor(Area_adj))) +
#   geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
#   scale_fill_viridis_d(name = "ha", na.value = "transparent", labels = leg) +
#   coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + thm()

leg <- c(nlabs, "")
p2 <- ggplot(africa) + 
  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
  geom_stars(data = nflds, aes(x = x, y = y, fill = factor(N))) +
  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
  scale_fill_viridis_d(name = "N", na.value = "transparent", labels = leg) +
  coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + thm()
# p2adj <- ggplot(africa) + 
#   geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
#   geom_stars(data = nflds_adj, aes(x = x, y = y, fill = factor(N_adj))) +
#   geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
#   scale_fill_viridis_d(name = "ha", na.value = "transparent", labels = leg) +
#   coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + thm()

op <- p + p2 + plot_annotation(tag_levels = 'A')

ggsave(op, bg = "transparent", width = 7, height = 4, units = "in", dpi = 300, 
       filename = pth("figures/si_fsize_n_05.png"))
```